### 2015-07-27 diss_titlepage.tex
Thesis title page

\thispagestyle{empty}
\vskip 3cm
\begin{center}
    {\Huge\bfseries\sf \mytitle \par}
    \vskip 1cm
    \sf by
    \vskip 1cm
    \large \sf \authorname \\
    \sf \authornumber
\end{center}

\begin{center}
	\large \sf Thesis presented in part-fulfilment of the degree of \degname \\ in accordance with the regulations of the \univname
\end{center}

\vskip 5cm

\begin{minipage}{0.4\textwidth}
\begin{flushleft} 
\large \sf 
\depname \\
\univname \\
\univaddr
\vskip 0.5cm
\copyright~\mydate~\authorname
\end{flushleft}
\end{minipage}

%\vskip 0.2cm

\begin{minipage}{\textwidth}
\begin{flushleft}
%\noindent
\sf This copy of the dissertation has been supplied on condition that anyone who consults it is understood to recognise that its copyright rests with the author and that no quotation from the dissertation, nor any information derived there from, may be published without the author's prior written consent. Moreover, it is supplied on the understanding that it represents an internal University document and that neither the University nor the author are responsible for the factual or interpretative correctness of the dissertation.
\end{flushleft}
\end{minipage}

\vfill

\newpage