### 2015-08-03 MSc_diss.tex
Updates in Neuman2012 review

\documentclass[11pt,a4paper]{article}
% Encoding and language
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
% Colors
\usepackage[usenames,dvipsnames]{color}
\usepackage{colortbl}
% Paper size and margins
%\usepackage{rotating}
\usepackage{pdflscape}
%\usepackage[a4paper,showframe]{geometry}
\usepackage{vmargin}
\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0cm}{0cm}{0cm}{1.5cm}
\usepackage{pdflscape}
\usepackage{afterpage}
% Indenting
\usepackage{indentfirst}
%\setlength{\parindent}{1cm}
\setlength{\parskip}{0.5cm}
% Symbols
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
% Equations
\usepackage{eulervm} % upright font in eqns (Palatino)
\usepackage{cool}
\usepackage{mathtools} % for arrows
\renewcommand{\theequation}{R\arabic{equation}}
% References
\usepackage{natbib}
\bibliographystyle{apalike}
% Hyperlinks and labels
%\usepackage{showkeys}
\usepackage[linktocpage=true,plainpages=false,pdfpagelabels=false]{hyperref}
\definecolor{citecolor}{rgb}{0.2,0.7,0.9}
\definecolor{urlcolor}{rgb}{0,0,1}
\hypersetup{
    colorlinks, linkcolor={Blue},
    citecolor={Blue}, urlcolor={Black}
}
% Figures
\usepackage{subcaption}
\usepackage{alphalph}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usepackage{epstopdf}
% Tikz styles
\tikzstyle{reg} = [rounded rectangle, text width=2cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
\tikzstyle{rad} = [rounded rectangle, draw=none, text width=2cm, fill=red!10, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
\tikzstyle{long} = [rounded rectangle, draw=none, text width=3.5cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]   
\tikzstyle{add} = [scale=1.5, draw=none,fill=none,align=center]
\tikzstyle{emp} = [draw=none,fill=none]
\tikzstyle{ar} = [->,draw=black!70,line width=2]
\tikzstyle{ln} = [-,draw=black!70,line width=2]
% Title etc
\newcommand{\mytitle}{Investigation of the relationship between tropospheric ozone production efficiency and carbon bond emissions}
\newcommand{\authorname}{Maria Zamyatina}
\newcommand{\authornumber}{Student No: 100106685}
\newcommand{\degname}{Master of Science}
\newcommand{\univname}{University of East Anglia}
\newcommand{\univaddr}{University Plain \\ Norwich NR47TJ UK}
\newcommand{\depname}{School of Environmental Sciences}
\newcommand{\mydate}{\the\year}
%% Remove author and date from the maketitle command
%\makeatletter
%\renewcommand{\@maketitle}{
%\newpage
% \null
% \vskip 2em%
% \begin{center}%
%  {\LARGE \@title \par}%
% \end{center}%
% \par} \makeatother

% PDF meta-data
\hypersetup{pdftitle={\mytitle}}
\hypersetup{pdfauthor=\authorname}

\title{\mytitle}
% =========================================================================
\begin{document}

\input{diss_titlepage}

\tableofcontents
\newpage

\begin{abstract}
Abstract.
\end{abstract}

\section{Introduction} \label{sec:intro}
The Earth’s atmosphere is the recipient of a vast range of chemical compounds emitted by natural and anthropogenic sources. The oxidation of these compounds is the main process taking place in the atmosphere that keeps their concentrations at levels that do not distort the chemical balance of the atmosphere \citep{Prinn2003}. In other words, oxidation is a removal 'cleansing' mechanism for these compounds (various pollutants and greenhouse gases), and its rate is often referred to as the oxidizing capacity of the atmosphere. This capacity depends mainly on two factors: climate and the global burden of the principal oxidants such as ozone ($O_3$), the hydroxyl radical ($OH$) and hydrogen peroxide ($H_2O_2$) \citep{Prinn2003,Thompson1992}.
\subsection{Tropospheric ozone}
Tropospheric ozone ($O_3$) plays an important role in the chemical and thermal balance of the atmosphere. Although it accounts only for about 10\% of the total ozone, it is a primary precursor to the formation of hydroxyl radicals ($OH$) which in association with hydrogen peroxide ($H_2O_2$) determines the 'cleansing' capacity of the atmosphere \citep{Prinn2003,Tarasick2008,Thompson1992}. On the other hand, being a source of main pollutant removal agents, ozone is an air pollutant itself. Long exposure to its high concentrations causes respiratory problems in humans and damage to sensitive plant species \citep{Fowler2008}. Last but not least characteristic of tropospheric ozone is that it is an important greenhouse gas with a current estimated radiative forcing of  $0.40\pm 0.20~Wm^{–2}$ which is approximately one fifth of the $CO_2$ radiative forcing \citep{Hartmann2013,Myhre2013}. Due to these reasons a considerable interest in quantifying tropospheric ozone concentrations, budget and trends exists for more than a 100 years \citep{Becker2004}.

Unlike many other air pollutants, tropospheric ozone is not directly emitted, but produced following the oxidation of carbon monoxide ($CO$), methane ($CH_4$), and nonmethane volatile compounds (nmVOCs) in the presence of nitrogen oxides (NOx) \citep{Crutzen1973,Myhre2013}. The emissions of these precursors arise from both natural and anthropogenic sources which have intensified since pre-industrial times \citep{Parrish2014,Volz1988} (Figure \ref{fig:O3observations}). Presumably it has led to roughly a doubling in background tropospheric ozone concentration \citep{Guicherit2000,Hartmann2013,Tarasick2008,Vingarzan2004}, although contribution from another major source, transport from the stratosphere, may also took place \citep{Fowler2008}. Climate is also believed to be a key driver of ozone production and destruction since the chemical reactions affecting ozone concentration are strongly dependent on such climatic factors as temperature, rainfall and humidity. Therefore in a changing climate tropospheric ozone is not anymore a local air quality issue, but is a global pollution problem \citep{Fowler2008}.

\begin{figure}[h]
\includegraphics[width=\linewidth]{{./pics/Isaksen2009_O3}.png}
\caption{Observed surface ozone at different Northern Hemispheric surface stations \citep{Isaksen2009}.}
\label{fig:O3observations}
\end{figure}

According to the Intergovernmental Panel on Climate Change (IPCC) Fifth Assessment report (AR5) for present conditions (around year 2000) the global mean tropospheric ozone budget is approximately 331 Tg \citep{Myhre2013}. However, despite general agreement on how the drivers impact ozone concentrations at global and regional scales, its budget varies considerably between different modelling and observational studies \citep{Stevenson2006,Wild2007,Young2012}. Table \ref{tab:O3budget} demonstrates that the range of the global tropospheric ozone budget estimates is approximately 60 Tg which is 18\% of the mean. In terms of ozone production the variation is even higher and accounts for 28\% of the corresponding value. In that regard, two questions arise: why does the tropospheric ozone budget differ between models and observations? And secondly, why does it vary to this extent, especially in terms of production? The only way to answer these questions is to perform a rigorous investigation of the factors that drive tropospheric ozone and attribute those differences to respective factors \citep{Young2012}.

MAKE GRAPHIC FROM TABLE ABOUT O3 BUDGETS!

\begin{table}[h] % O3 budget
\centering
\caption{Summary of tropospheric ozone global budget model and observation estimates for present (about 2000) conditions. All uncertainties quoted as 1 standard deviation (68\% confidence interval). Adapted from \citep{Myhre2013}}
\label{tab:O3budget}
\begin{tabular}{ccl}
\hline
Burden, Tg & Production, Tg yr–1 & Reference \\
\hline
\multicolumn{3}{c}{Modelling studies} \\
$337\pm23$  & $4877\pm853$ & Young et al. (2013); ACCMIP \\
$323$	    & -	           & Archibald et al. (2011) \\
$330$	    & $4876$       & Kawase et al. (2011) \\
$312$       & $4289$       & Huijnen et al. (2010) \\
$334$       & $3826$       & Zeng et al. (2010) \\
$324$       & $4870$       & Wild and Palmer (2008) \\
$314$       & -	           & Zeng et al. (2008) \\
$319$	    & $4487$	   & Wu et al. (2007) \\
$372$	    & $5042$       & Horowitz (2006) \\
$349$	    & $4384$	   & Liao et al. (2006) \\
$344\pm39$	& $5110\pm606$ & Stevenson et al. (2006); ACCENT \\
$314\pm33$	& $4465\pm514$ & Wild (2007) (post-2000 studies) \\
\hline
\multicolumn{3}{c}{Observational studies} \\
$333$   	& -	           & Fortuin and Kelder (1998) \\
$327$	    & -	           & Logan (1999) \\
$325$	    & -	           & Ziemke et al. (2011); 60S–60N \\
$319–351$	& -	           & Osterman et al. (2008); 60S–60N \\
\hline
\end{tabular}
\end{table}

A considerable effort has been made to tackle these problems, and generally there are two possible ways of doing it. The first one implies the direct analysis of tropospheric chemical schemes used in climate-chemistry models. For example, Emmerson and Evans (2009) compared the state of the art Master Chemical Mechanism which contains approximately 5600 species and 13500 reactions \citep{Jenkin2002} with other six chemical schemes and found four significant variations between them. This approach is linked with big challenges since a modern global model is a very sophisticated representation of the climate-chemistry interactions, and therefore it is often difficult to separate influence of one reaction from another. However, even being so developed climate-chemistry models do not take into account all known chemical reactions taking place in the real troposphere. A complete representation requires many thousands of species and tens of thousands of reactions, which is beyond the numerical capabilities currently available. This significantly limits the practicable size of chemical schemes and typically involves a reduction of the number of VOCs considered and lumping the carbon from the discarded species into representative surrogates.

! add something from report!

The second way to understanding the discrepancies between models and observations is to conduct an idealised theoretical study using a conceptual model. The advantage of this approach is that such kind of models can be quite easily constructed, and more importantly can be focused on representation of a certain chemical reactions which presumably could explain the discrepancy between models and observations.

\subsubsection*{Ozone tropospheric chemistry}
%\input{O3CycleScheme}
Having identified the role of tropospheric ozone, it is time to remind ourselves the chemistry that controls abundance of it.

The key chemical processes that lead to $O_3$ formation and destruction in the troposphere are driven by reaction cycles involving free-radical intermediates, which are mainly formed from the photolysis of $O_3$ itself (\ref{reac:O3+hv=O1D+O2}) \citep{Fowler2008}. At UV wavelengths shorter than 320 nm, $O_3$ photolysis generates electronically excited oxygen atoms ($O(^1D)$), which than can either react with water vapour to form $OH$ radicals (\ref{reac:O1D+H2O=OH+OH}), or collide with an inert molecule (denoted '$M$') to reform $O_3$ (\ref{reac:O1D=O3P} and \ref{reac:O3P+O2=O3}):
\begin{equation}\label{reac:O3+hv=O1D+O2}
O_3 + hv \rightarrow O(^1D) + O_2
\end{equation}
\begin{equation}\label{reac:O1D+H2O=OH+OH}
O(^1D) + H_2O \rightarrow OH + OH
\end{equation}
\begin{equation}\label{reac:O1D=O3P}
O(^1D) \xrightarrow{M} O(^3P)
\end{equation}
\begin{equation}\label{reac:O3P+O2=O3}
O(^3P) + O_2 \xrightarrow{M} O_3
\end{equation}
Once formed the OH radicals mainly react with $CO$ and $CH_4$ and initiate the $O_3$ formation or removal cycles by producing the peroxy radicals ($HO_2$, $CH_3O_2$):
\begin{equation}\label{reac:OH+CO=H+CO2}
OH + CO \rightarrow H + CO_2
\end{equation}
\begin{equation}\label{reac:H+O2=HO2}
H + O_2 \xrightarrow{M} HO_2
\end{equation}
\begin{equation}\label{reac:OH+CH4=CH3+H2O}
OH + CH_4 \rightarrow CH_3 + H_2O
\end{equation}
\begin{equation}\label{reac:CH3O+O2=CH3O2}
CH_3 + O_2 \xrightarrow{M} CH_3O_2
\end{equation}
The fate of the peroxy radicals depends on the presence of $NO_x$ ($=NO + NO_2$). In low-$NO_x$ conditions (remote regions), the $HO_2$ radical reacts with another $HO_2$ to form hydrogen peroxide ($H_2O_2$) or reacts with the methyl peroxy radical ($CH_3O_2$) to form methyl hydrogen peroxide ($CH_3OOH$):
\begin{equation} \label{reac:HO2+HO2=H2O2+O2}
HO_2 + HO_2 \rightarrow H_2O_2 + O_2
\end{equation}
\begin{equation} \label{reac:HO2+CH3O2=CH3OOH+O2}
HO_2 + CH_3O_2 \rightarrow CH_3OOH + O_2
\end{equation}
On a whole, in low-$NO_x$ conditions, a net $O_3$ destruction takes place, because the reaction sequence is initiated by $O_3$ photolysis. Plus in the same conditions some additional $O_3$ removal can occur because of the reaction of the $HO_2$ radical with $O_3$ leading to further destruction of ozone in a chain sequence involving formation of $OH$ radicals:
\begin{equation}\label{reac:HO2+O3=OH+2O2}
HO_2 + O_3 \rightarrow OH + 2O_2
\end{equation}
\begin{equation}\label{reac:OH+O3=HO2+O2}
OH + O_3 \rightarrow HO_2 + O_2
\end{equation}

In intermediate $NO_x$ conditions (rural areas of most industrialised countries), the peroxy radicals react with nitrogen monoxide ($NO$) to form nitrogen dioxide ($NO_2$), which subsequent photolysis leads to $O_3$ generation:
\begin{equation}\label{reac:HO2+NO=NO2+OH}
HO_2 + NO \rightarrow NO_2 + OH
\end{equation}
\begin{equation}\label{reac:CH3O2+NO=NO2+CH3O}
CH_3O_2 + NO \rightarrow NO_2 + CH_3O
\end{equation}
\begin{equation}\label{reac:NO2+hv=NO+O3P}
NO_2 + hv \rightarrow NO + O(^3P)
\end{equation}
\begin{equation}\label{reac:O3P+O2=O3}
!!!(R4)sameO(^3P) + O_2 \xrightarrow{M} O_3
\end{equation}
As shown in Figure ??? reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:CH3O2+NO=NO2+CH3O}) form parts of the free-radical propagated $O_3$ forming cycles, which may occur a number of times before being halted by radical termination reactions (\ref{reac:HO2+HO2=H2O2+O2} and \ref{reac:HO2+CH3O2=CH3OOH+O2}), and the higher $NO_x$ level is the greater number of these cycles can occur. Therefore, in intermediate $NO_x$ conditions a net $O_3$ formation takes place and it is sensitive to $NO_x$.

In high-$NO_x$ conditions (urban environment), the $OH$ radical may have another fate rather than reacting with $CO$, $CH_4$ and $O_3$. Depending on the concentration of $NO_2$ it can react with it to form nitric acid ($HNO_3$):
\begin{equation}\label{reac:OH+NO2=HNO3}
OH + NO_2 \xrightarrow{M} HNO_3
\end{equation}
The formation of $HNO_3$ terminates free-radical propagated $O_3$ forming cycles leading to a decrease in $O_3$ production rate. However, elevated emissions of $CO$ and $CH_4$ (and other nonmethane volatile organic compounds (nmVOCs)) may make reactions (\ref{reac:OH+CO=H+CO2})-(\ref{reac:OH+CH4=CH3+H2O}) more competitive with the reaction (\ref{reac:OH+NO2=HNO3}) and lead to an increase in the $O_3$ production rate. Therefore, in high-$NO_x$ conditions the $O_3$ production rate can either decrease or increase depending on $NO_x$ and $VOCs$ levels.

The free-radical propagated oxidation of hydrocarbons leads to the generation of carbonyl compounds, among which formaldehyde ($HCHO$) is the most common. Its photolysis and further oxidation contributes to radical generation:
\begin{equation}\label{reac:HCHO+hv=H+HCO}
HCHO + hv \rightarrow H + HCO
\end{equation}
\begin{equation}\label{reac:HCO+O2=HO2+CO}
HCO + O_2 \rightarrow HO_2 + CO
\end{equation}
\begin{equation}\label{reac:H+O2=HO2}
!!!sameH + O_2 \xrightarrow{M} HO_2
\end{equation}
therefore also having an impact on the rates of $O_3$ production \citep{Fowler2008} as do some other species, which are not mentioned in this section. Their role is described in other sections as far as it is necessary.

\subsection{Alkyl nitrates}
Alkyl nitrates ($RONO_2$) are organic trace gases, which through one of their formation mechanisms are closely linked with the production of tropospheric ozone \citep{Reeves2007}. This mechanism is present in the following section.
\subsubsection*{Alkyl nitrates chemistry}
The photochemical formation of the alkyl nitrates starts when chemistry of the troposphere reaches the regime of $O_3$ production, i.e. when reactions of the peroxy radicals with $NO$ start to compete with their self- and cross-reactions more effectively due to an increase in $NO_x$ mixing ratio. In this case the reaction between the alkyl peroxy radical ($RO_2$, where $R$ denotes methyl group, $CH_3$) and $NO$ can reveal its two-channel nature \citep{Day2003}:
\begin{subequations} \label{reac:RONO2alpha}
\begin{align}
RO_2 + NO &\xrightarrow{1-\alpha_3} RO + NO_2 \label{reac:RO+NO=aphaRO+NO2}\\
RO_2 + NO &\xrightarrow{\alpha_3} RONO_2 \label{reac:RO+NO=aphaRONO2}
\end{align}
\end{subequations}
The latter reaction can serve as a termination step of the propagated $O_3$ forming cycle (Figure ???), and as \citep{Roberts1998} estimated it is likely to begin to occur when the $NO$ mixing ratio increases up to 100-200 ppt. The yield of alkyl nitrate formation is controlled by the branching ratio ($\alpha_3$), which is temperature and pressure dependent and increases with the number of carbons in the alkyl peroxy radical \citep{Roberts1990}.

An alternative photochemical pathway of the formation of alkyl nitrates is the reaction of the alkoxy radical ($RO$) with $NO_2$:
\begin{equation}\label{reac:RO+NO2=RONO2}
RO + NO_2 \rightarrow RONO_2
\end{equation}
This pathway has been shown to become dominant over reaction (\ref{reac:RO+NO=aphaRONO2}) only for the production of lower (methyl ($C_1$), ethyl ($C_2$)) nitrates when $NO_x$ mixing ratio exceeds 35 ppb \citep{Archibald2007}, i.e. in highly polluted environments.

Alkyl nitrates are removed from the atmosphere by photolysis \citep{Turberg1990} and $OH$ radical oxidation \citep{Talukdar1997}:
\begin{equation} \label{reac:RONO2+hv=RO+NO2}
RONO_2 + hv \rightarrow RO + NO_2
\end{equation}
\begin{equation} \label{reac:RONO2+OH=RCHO+NO2}
RONO_2 + OH \rightarrow RCHO + NO_2
\end{equation}
The relative importance of these two loss reactions depends generally on the net UV intensity and the size of the carbon chain in the alkyl nitrate. Photolysis is relatively more important for nitrates with $< C_4$, while nitrates with longer chains ($\geq C_5$) tend to react faster with $OH$ \citep{Worton2010}.

The lifetimes of alkyl nitrates depend on the seasonality of the photolysis and $OH$ sink, which in turn vary with latitude and altitude. According to \citep{Clemitshaw1997} and \citep{Turberg1990} alkyl nitrates have atmospheric lifetimes of several days to several years, which makes them temporary reservoirs for the short lived $NO_x$ compounds. Being transported from polluted areas to remote background regions and eventually destroyed alkyl nitrates can 'release' $NO_2$ back to the atmosphere, thereby contributing to tropospheric ozone formation.
\subsection{Ozone, alkyl nitrates and carbon bonds}
In the previous sections we covered the chemistry that controls the production of $O_3$ in the troposphere. A traditional way to assess the efficiency of this production is to calculate the ratio:
\begin{equation}\label{eq:OPE}
OPE = \dfrac{k_{HO_2+NO}[HO_2][NO]}{k_{OH+NO_2}[OH][NO_2]} = \dfrac{P_{O_3}}{L_{NO_x}}
\end{equation}
where $OPE$ is the ozone production efficiency, $k_{HO_2+NO}$ - the rate coefficient of reaction (\ref{reac:HO2+NO=NO2+OH}), $k_{OH+NO_2}$ - the rate coefficient of reaction (\ref{reac:OH+NO2=HNO3}), $[X]$ - mixing ratios of corresponding species, $P_{O_3}$ – the $O_3$ production rate, $L_{NO_x}$ – the $NO_x$ loss rate. $OPE$ is determined by the amount of $HO_2$ reacting with $NO$ to form $NO_2$ and $OH$ (i.e., $k_{HO_2+NO}[HO_2][NO]$, (\ref{reac:HO2+NO=NO2+OH})) and the amount of $NO_x$ removed from the system by reaction of $OH$ with $NO_2$ to form nitric acid ($HNO_3$, reaction (\ref{reac:OH+NO2=HNO3})). In other words, $OPE$ shows the amount of ozone created ($P_{O_3}$) per molecule of $NO_x$ removed ($L_{NO_x}$) from the system \citep{Jacob1999}. Due to the reasons described in section \ref{sec:chem_mech} (comment at the bottom of the page!"No representation of the nitric acid ($HNO_3$) in the model"), we use another expression for calculations of the $O_3$ production efficiency. This expression has been recently proposed by Professor Matthew Evans (University of York) amid the mounting concerns about the need of thorough intercomparison of chemical mechanisms of global atmospheric models (including those used in the IPCC reports), why it is more universal and applicable to any kind of model.

According to \citep{Evans2014} (personal communication), the $O_3$ production efficiency can be estimated via calculation of the number of carbon bonds emitted into the system. For example, if the system contains one molecule of methane, $CH_4$, as 'a carbon bond reservoir', it means that this system has 4 carbon bonds (4 C-H), if  instead of methane it contains one molecule of ethane, $C_2H_6$, the system has 7 carbon bonds (6 C-H bonds and 1 C-C), and so on \citep{Edwards2013}. Assuming that every carbon bond is broken chemically and every organic peroxy radical ($RO_2$) reacts with $NO$, the overall efficiency of carbon bonds conversion to $O_3$, $\epsilon$, can be calculated as: 
\begin{equation}\label{eq:epsilon}
\epsilon = \dfrac{P_{O_3}}{E_{C_{bons}}} = \chi\cdot\upsilon
\end{equation}
where $P_{O_3}$ – the rate of $O_3$ production (mol yr-1), $E_{C_{bons}}$  – the rate of carbon bonds emission (mol yr-1), $\chi$ – ratio between the rate of carbon bonds loss to form $RO_2$ and the loss rate of all carbon bonds, $\upsilon$ – ratio between the rate of $RO_2$ loss in reaction with $NO$ resulting in $NO$-to-$NO_2$ conversion and the loss rate of all $RO_2$. In other words, chemical loss of 1 carbon bond produces 1 peroxy radical:
\begin{itemize}
\centering
\item[] $OH + {CH_4}|_{4 bonds} \rightarrow {CH_3O_2}|_{3 bonds} + H_2O$
\item[] 4 carbon bonds = 3 carbon bonds, 1 peroxy radical
\end{itemize}
Loss of 1 carbon bond via photolysis produces 2 peroxy radicals:
\begin{itemize}
\centering
\item[] ${HCHO}|_{2bonds} + hv \rightarrow H + {HCO}|_{1bond}$
\item[] ${HCO}|_{1bond} + O_2 \rightarrow HO_2 + {CO}|_{1bond}$
\item[] $H + O_2 \xrightarrow{M} HO_2$
\item[] $horizontal line$
\item[] 2 carbon bonds = 1 carbon bond, 2 peroxy radicals
\end{itemize}
or no peroxy radicals because of dry or wet deposition:
\begin{itemize}
\centering
\item[] $HCHO \xrightarrow{deposition}$
\item[] 2 carbon bonds = 2 carbon bonds, 0 peroxy radicals
\end{itemize}

The fact that $\epsilon$ does not account for reactions where no $NO$-to-$NO_2$ conversion happens, as, for example, in case of the formation of the alkyl nitrates, $\epsilon$ provides a way to relate the rate of $O_3$ production to the alkyl nitrate chemistry. Knowing the rate of carbon bond emissions ($E_{C_{bons}}$) from emissions inventory data and knowing the rate of $O_3$ production from a model run ($P_{O_3}$), one can estimate the efficiency of $O_3$ production per carbon bond emitted in their model, and than compare it with other such values from different models potentially revealing the true cause of discrepancies between models and observations of tropospheric $O_3$ budgets.

For link at the bottom? The same kind of approach is used in the Common Representative Intermediates (CRI) mechanism for VOC degradation \citep{Jenkin2002, Jenkin2008}. It is a reduced chemical mechanism that describes $O_3$ formation from the tropospheric degradation of methane and 115 emitted non-methane hydrocarbons and oxygenated volatile organic compounds developed using the Master Chemical Mechanism version 3.1 (MCM v3.1) as a reference benchmark
\section{Overall objective and specific aims}
As the reader might anticipateit is mentioned in section \ref{sec:lit_review}
\citep{Sommariva2008}
[Typically, chemical mechanisms do not contain
enough information about the various formation routes of secondary organic compounds, leading to the underestimation of their concentrations, with possible consequences on the estimates of ozone production in remote regions, or to ineffective ozone-control policies. In this work, a quasi-explicit chemical model based upon the Master Chemical Mechanism (MCM) has been used to study in detail the photochemistry of a number of alkyl nitrates and determine quantitatively their formation processes. This kind of study shows the potential of using mechanisms with a high level of chemical detail to study the formation of pollutants in the atmosphere (Madronich, 2006).]
\subsection{Overall objective}
The overall purpose of this research is to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
\subsection{Specific aims}
Specific aims are to investigate:
\begin{itemize}
\item[(1)] the relationship between ozone production and carbon bond emissions;
\item[(2)] the impact of inclusion of alkyl nitrates chemistry on this relationship;
\item[(3)] how (1) and (2) vary in different $NO_x$ and $VOC$ regimes;
\item[(4)] build different version of the box model;
\item[(5)] to evaluate the results.
\end{itemize}

\section{Literature review} \label{sec:lit_review}
A lot of work is has been dedicated to measurements of alkyl nitrates in different environments and even more to measurements of ozone. The majority of these studies have been done in a observational-modelling framework.

During/After one of these investigations, namely the SAGA campaign, \citep{Atlas1993} proposed by that the short-chained alkyl nitrates ($C_1$-$C_3$) can be produced not only photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990}, but also emitted directly from equatorial oceans. This has since been
confirmed by field measurements \citep{Chuck2002,Blake2003,Dalh2005}.

There have been a number of studies of the latitudial and seasonal variations of alkyl nitrates in marine air (Blake et al., 2003), in the Arctic (Swanson et al., 2003) and in the
Antarctic (Beyersdorf et al., 2010)
\subsubsection{Observational studies}
Anthropogenic!!! Neuman
Source characterization and photochemical processes in transported air masses have been analyzed by examining RONO2 formation from anthropogenic emissions of alkane precursors [Simpson et al., 2003; Madronich, 2006; Simpson et al., 2006; Reeves et al., 2007; Worton
et al., 2010]

\citep{Sommariva2008}
[Alkyl nitrates typically constitute only a small
fraction of total NOy <10\%, even though they can be a much larger component in the equatorial marine boundary layer (Blake et al., 2003; Simpson et al., 2006 and references therein). Due to their long lifetime (41 week) alkyl nitrates can transport NOx to remote regions, with potential impacts on regional O3 formation (e.g., Murphy et al., 2006; Simpson et al., 2006)]

\subsubsection*{Continent}
\citep{Day2003}
We describe measurements of NO2, total peroxy nitrates (?PNs), total alkyl nitrates (?ANs), and HNO3 using thermal dissociation followed by laser-induced fluorescence detection of NO2 at three continental locations. At a rural site in California, measurements over a full annual cycle show that ?ANs are routinely 10–20\% of NOy. rural site, at a suburban site in California and an urban site in Houston, Texas [rural (UC-BFRS), suburban (Granite Bay) and urban (La Porte)]
SumANs a more broad term than usual.
?AN mixing ratios at the UC-BFRS exhibit a strong seasonal cycle with a minimum in winter and a maximum in summer. The seasonal transition is sharp. Concentrations of ?ANs rise rapidly from 150 ppt at the end of April to 500 ppt by early June.
We also show that ?ANs are a large part, if not all, of the ‘‘missing NOy’’.

\citep{Simpson2006} Pearl River Delta
Mixing ratios of seven C1–C5 alkyl nitrates (RONO2) were measured during a 16-month study (August 2001–December
2002) at Tai O, a coastal site 30km west of central Hong Kong in the Pearl River Delta, the fastest-growing industrial region in the world. The C3–C4 (rather than C1–C2) RONO2 were most abundant throughout the study, showing the importance of photochemical (rather than marine) RONO2 production in the sampled air. A lack of methyl nitrate (MeONO2) enhancement during summer, when the prevailing winds are from the ocean, indicates that the South China Sea is not a region of strong RONO2 emissions. By contrast, MeONO2 levels during pollution episodes (up to 25 parts per trillion by volume (pptv)) were the highest that our group has recorded during urban photochemical RONO2 production, as opposed to marine emissions or biomass burning. The highest summed RONO2 level of the study (204 pptv) was measured in the afternoon of 7 November 2002, during an intense pollution episode that captured the highest ozone (O3) level ever recorded in Hong Kong (203 ppbv). During pollution episodes, the average ratio of O3 to summed RONO2 was
roughly 1000:1 in freshly polluted air (ethyne/CO?3–5 pptv/ppbv) and 500:1 in very freshly polluted air (ethyne/ CO?6–8 pptv/ppbv). Ozone and RONO2 share a common photochemical source, and their good correlation in pollution plumes shows that RONO2 can be used as a tracer of photochemical O3 production. Even MeONO2 showed similar
diurnal variations as the C2–C5 RONO2, indicating a strong photochemical source despite its very slow photochemical production from methane oxidation. The decomposition of longer-chain alkoxy radicals also does not explain the high MeONO2 levels, and rough calculations show that methoxy radical reaction with NO2 appears to be a viable alternate pathway for MeONO2 production in polluted atmospheres, though further measurements and modeling are required to confirm this mechanism.
\citep{Comment}
It has generally been assumed that reaction (3) dominates over reaction (2) and that production of CH3ONO2 would be small. Simpson et al. (2006) suggest that model studies would be useful to test their theory. In this short comment on their paper, some model simulations are presented that confirm their hypothesis.The model simulations have been carried out using a box model running the CRI v1.0 mechanism (Jenkin et al., 2002), which is a reduced scheme of the Master Chemical Mechanism (Jenkin et al., 1997). The box model was integrated using emissions of NOx and VOC taken from the UK National emissions inventory and would be repre- sentative therefore of European rather than Asian emission profiles but are still instructive. The model is designed to simulate regional photochemical episodes and runs during European summer time (mid June) under moderately high temperature (approximately 25 1C) and relative humidity (70\%). The concentration of CH3ONO2 from the two routes, reaction (1b) and reaction (2), have been compared as a function of NOx and are summarised in Fig. 1. It is clear that above about 10 ppb of NO2 channel (2) becomes important as a source of CH3ONO2 and at around 35 ppb starts to become the dominant channel in the simulations performed.

\citep{Stroud2001} plane+colorado, plane+North Atlantic
During both campaigns alkyl nitrates were measured aboard the NOAA Orion WP-3 aircraft with a vertical range from the boundary layer to the midtroposphere (7 km) [Hiibler et al.,1998]. During STERAO 1996 the aircraft was based at Buckley Air Force Base in Denver, Colorado. Flights were conducted from June 13, 1996, to July 16, 1996, over the northeastern plains of Colorado to study how thunderstorms affect the chemical composition of the atmosphere.
Alkyl nitrates were measured over the northeastern plains of Colorado during the Stratospheric Tropospheric Exchange: Radiation, Aerosols and Ozone (STERAO) 1996 campaign and over the North Atlantic during the North Atlantic Regional Experiment (NARE) 1997. A comparison of alkyl nitrate to hydrocarbon ratios with a sequential reaction scheme model suggests our understanding of C4-Cs alkyl nitrate photochemistry is adequate to reproduce the observations. Air masses with larger ratios of alkyl nitrate to hydrocarbons were observed during NARE 1997 than STERAO 1996, consistent with the NARE 1997 measurements being in the more remote North Atlantic. These results suggest that information related to the photochemical age of an air  mass can be derived from alkyl nitrate to hydrocarbon ratios.

\citep{Worton2010} TORCH
Alkyl nitrates (C1eC5) were measured at two sites (near urban and rural) in southeast England during the Tropospheric Organic Chemistry Experiment (TORCH). Methyl nitrate was the dominant species during both campaigns accounting for on average about one third of the total measured alkyl nitrates. High mixing ratios (>50 pptv) and variability of methyl nitrate were observed at the near urban site (TORCH1) that were not seen at the rural site (TORCH2) and which could not be explained by local photochemical production or direct emissions. The diurnal variation of methyl nitrate during TORCH1 showed a morning maximum that would be consistent with nighttime chemistry followed by transport to the surface by boundary layer dynamics. Similarly, elevated morning mixing ratios were also observed during TORCH2 although the magnitudes were much smaller. As a result, methyl nitrate could represent a tracer for nighttime chemistry seen at the ground the following day. At both campaigns, the dominant source of short chain alkyl nitrates and carbonyl precursor radicals (?C4) were from decomposition of larger compounds. The magnitude of the source increased with decreasing carbon number consistent with increasing total precursor abundance. Non-photochemical emissions of acetaldehyde and acetone could not be accounted for by automobile exhaust emissions alone and indicated that other direct sources are likely important in this environment.
Alkyl nitrate measurements were made at two sites (near urban
and rural) as part of the TORCH experiment. Observed mixing ratios
were higher (factor of ?2) during TORCH1, especially during the identified heatwave period. Methyl nitrate accounted for w30\% of the total measured C1eC5 species and during TORCH1 was often present at mixing ratios > 50 pptv. The diurnal profile of methyl nitrate showed an early morning maximum that we attributed to nighttime production from NO3 chemistry in the residual layer followed by transportation to the surface by boundary layer dynamics. Further work is needed to evaluate our hypothesis of a nighttime source of methyl nitrate but if proven, methyl nitrate could represent an interesting tracer of nighttime chemistry aloft. The predominant photochemical source of the short chain alkyl nitrates and carbonyl compounds was from the photochemical oxidation and decomposition of longer chain compounds as opposed to the oxidation of the parent hydrocarbons similar to previous studies. The magnitude of this source shows a preference for the shortest chain species consistent with the increasing number of precursor compounds identified in the MCM. Primary sources of acetaldehyde and acetone were also significant during TORCH1 and could not be accounted for by automobile exhaust emissions alone. Further work to identify and quantify emissions from additional direct sources is necessary if appropriate control strategies are to be implemented.

\citep{deKock1994}
Air samples were collected at Cape Recife (34 1' S 25 40'E) situated on the south-east coast of Southern Afi'ica (Figure 1). The suburban, urban and industrial areas of Port Elizabeth lie approximately 10-40 kin north to north-west of the sampling site.
Mean sumANs 12.7.
Variations in absolute and relative concentrations of the alkyl nitrates could be related to changes in meteorological conditions. It is considered that the differences observed in chemical composition may represent different sources of trace gas emissions or different degrees of photochemical processing of similar sources. The meas~ed levels of the alkyl nitrates suggest long-range transport to play a major role in regulating the distribution of alkyl nitrates in the troposphere.

\citep{Beyersdorf2010}
South Pole, other Antarctic, ANs only

\subsubsection*{Oceanic sources}
\citep{Newland2013}
[An oceanic source for C1 – C3 alkyl nitrates was first postulated by Atlas et al. (1993) during the SAGA campaign in which it was observed that elevated mixing ratios of methyl and ethyl nitrates correlated well with equatorial regions of high productivity. This has since been confirmed by field measurements (e.g. Chuck et al., 2002; Blake et al., 2003; Dahl et al., 2005). Chuck et al. (2002) estimate an annual emission flux from equatorial Atlantic and Pacific Ocean regions of 230 and 68 Gg yr-1 for methyl and ethyl nitrate respectively. The oceanic source of more or equal C4 alkyl nitrates is insignificant compared to anthropogenic sources (Blake et al., 2003).
The production mechanism of the alkyl nitrates is thought to be by the reaction ROO + NO in the aqueous phase with the peroxy radical coming from the photolysis of coloured dissolved organic matter (CDOM) and the NO coming from the photolysis of nitrite (Dahl et al., 2003). Direct emission from algae (Chuck et al., 2002) or from bacteria at depth (Dahl et al., 2007) has also been postulated.]

\citep{Atlas1993}
Alkyl Nitrates, Nonmethane Hydrocarbons, and Halocarbon Gases Over the Equatorial Pacific Ocean During Saga 3
February March 1990
The cruise covered a region of the equatorial Pacific Ocean from 15N to 10S latitude and 144 to 165W longitude. On this cruise we collected samples for the measurement of alkyl nitrates (RONO2) (C2-C5), nonmethane hydrocarbons (NMHC) and several halocarbon gases.
Results:
- Alkyl nitrates were found at levels higher than predicted from steady state calculations based on measured mixing ratios of hydrocarbons and NO. The measured levels of RONO2 suggest long-range transport as one mechanism contributing to elevated concentrations of alkyl nitrates in the remote troposphere. However, the distributions of C2 and C3 alkyl nitrates over the equator were  similar to the organobromine gases. This distribution suggests
a possible oceanic source for alkyl nitrates to the atmosphere.
Model: 1D 41 constituents incl. 3 ANs

\citep{Chuck2002} September, October
Measurements of methyl and ethyl nitrate in seawater and air samples along two Atlantic Ocean transects provide the FIRST direct evidence for an oceanic source of these compounds. A simple box model calculation suggests that the equatorial source could be a major component of the local atmospheric alkyl nitrate budget.
In the remote atmosphere, the concentration of methyl nitrate (MeONO2) is expected to be low because of low concentrations of NOx and a branching ratio of less than 1\% for the formation of MeONO2
from the reaction of alkylperoxy radicals with NO (8). However, unexpect- edly high levels of RONO2 in marine air and observations of a strong positive correlation between ethyl and propyl nitrate and the marine biogenically produced halocarbon, bromoform, in the marine boundary layer over the equatorial Pacific Ocean led Atlas et al. to propose an oceanic source for alkyl nitrates in the atmo- sphere (9). Subsequent atmospheric sampling campaigns, mostly conducted over the remote Pacific Ocean, have provided yet more evidence to support the hypothesis of an oceanic source (2, 10), with the contribution of alkyl nitrates to the NOy
sum reaching 80\% in equatorial and high-latitude regions over the South Pacific (Blake, 1999). Although there have been no pub- lished measurements confirming the oceanic emission of these species, other studies have also invoked a seawater source for methyl, eth- yl, and propyl nitrates (11, 12). The supersaturation of surface waters over large areas of the Atlantic provides the first direct evidence for an oceanic source of MeONO2 and EtONO2. dominant source region is found between the equator and 40S. distinct mechanisms may be respon- sible for the production of methyl and ethyl nitrate, with different water types supporting different mechanisms. Methyl nitrate generally showed profiles more typically associated with algal production, with decreasing concentrations be- low the chlorophyll maximum. In contrast, in waters more strongly affected by continental influences, such as increased loading of dis- solved organic matter, the alkyl nitrates showed maximum concentrations in surface waters, which may suggest that photochemical production is important in shelf and coastal regions.

\citep{Blake2003} oceanic+urban!!!
Observations: concentration distributions of C1-C4 alkyl nitrates observed during the NASA airborne campaigns Pacific Exploratory Mission (PEM) -Tropics A (September– October 1996) and PEM-Tropics B (March–April 1999). The total geographic range for PEM-Tropics A was 45?N–72?S latitude and 153?E–75?W longitude, and for PEM- Tropics B was 40?N–36?S latitude and 149?E–75?W longitude. We observed high methyl nitrate (MeONO2,CH3ONO2) mixing ratios (approximately 50 pptv) at low altitudes in a latitude band between 8?Nto13?S stretching across the equatorial Pacific, illustrating the oceanic source of MeONO2. This source may be associated with the high-nutrient, low-chlorophyll character of equatorial Pacific waters. We discuss MeONO2 and ethyl nitrate (EtONO2,C2H5ONO2), whose abundance is dominated by equatorial oceanic sources, 2-Propyl nitrate (2-PrONO2,2-C3H7ONO2), which has significant oceanic and northern hemispheric (NH) sources associated with urban/ industrial hydrocarbon emissions, and 2-butyl nitrate (2-BuONO2 2-C4H8ONO2), which has mostly NH (northern hemisphere) sources.
the ocean as the major source for MeONO2
Ethyl nitrate has a dominant oceanic source in equatorial regions, but at mid NH latitudes a small additional source is associated with urban/industrial emissions.

\citep{Reeves2007} ocean+urban+fire
This paper is based on alkyl nitrate measurements made over the North Atlantic as part of the International Consortium for Research on Atmospheric Transport and Transformation (ICARTT). The focus is on the analysis of air samples collected on the UK BAe-146 aircraft during the Intercontinental Transport of Ozone and Precursors (ITOP) project, but air samples collected on board the NASA DC-8 and NOAA WP-3D aircraft as part of a Lagrangian experiment are also used. The ratios between the alkyl nitrates and their parent hydrocarbons are compared with those expected from chemical theory. Further, a box model is run to investigate the temporal evolution of the alkyl nitrates in three Lagrangian case studies and compared to observations.
The general agreement between the alkyl nitrate data and photochemical theory suggests that during the first few days of transport from the source region, photochemical production of alkyl nitrates, and thus ozone, had taken place.
Concentrations of methyl, ethyl, 1-propyl, 2-propyl, 2-butyl, 2-methyl-3-butyl and the sum of 2- and 3-pentyl nitrates (hereafter referred to as 2 + 3-pentyl nitrate) are presented.
The alkyl nitrate concentrations observed during
ITOP are typical of those found a few days downwind of precursor source regions. Although there is some evidence for an oceanic source of methyl nitrate, for the most part the air samples collected during ITOP do not appear to be strongly influenced by oceanic sources of the alkyl nitrates. Elevated concentrations were observed in air masses which have been influenced by emissions from the N.E. United States and from Alaskan fires. The large variations in ratios between the C5 and C2-C4 nitrates suggest that emission ratios for the parent hydrocarbons vary between different regions of the N.E. United States. [49] In general the alkyl nitrate data agrees well with photochemical theory. This suggests that at least during the first few days of transport from the source region photo- chemical production of alkyl nitrates had taken place. This implies that ozone production had also occurred. There also appears to be a widespread common source of ethyl nitrate and 1-propyl nitrate other than from their parent hydro- carbons. This might be from longer-chain hydrocarbons, possibly pentanes. [50] The observations in the more photochemically pro- cessed air masses are consistent with the concentration of NO dropping to a point where the rates of the alkyl nitrate production reactions no longer dominate the peroxy radical self/cross reactions. However, in these low levels of NO, where the balance between ozone production and destruc- tion is highly sensitive to NOX, the alkyl nitrates might contribute to maintaining the ozone levels in air as it undergoes long-range transport. [51] The observed patterns of the alkyl nitrates also suggest that the rates of photochemical processing in the Alaskan smoke plumes were small and that the rates of ozone formation in these air masses were low.

\citep{Simpson2003} airborne too
TRACE-P samples that did not show influence from marine alkyl nitrate sources were used to investigate photochemical alkyl nitrate evolution. Relative to 2-butyl nitrate/n-butane, the measured ratios of ethyl nitrate/ethane and 2-propyl nitrate/propane showed notable deviations from modeled values based on laboratory kinetic data, suggesting additional Asian sources of their alkyl peroxy radical precursors. By contrast, the measured ratios of 1-propyl-, 2-pentyl-, and 3-pentyl nitrate to their respective parent hydrocarbons were fairly close to modeled values.

\citep{Roberts1998} marine MBL+urban ANs
The measurements described herein were made as part of the
Organic nitrates are a class of product species that contain
North Atlantic Regional Experiment (NARE) during August and September 1993. These measurements consisted of NO, NO 2, particulate nitrate, HNO3, PAN, PPN, RONO 2, NOy, 0 3, CO, SO 2, and C2-C6 hydrocarbons at Chebogue Point. This site was ideally located to study the evolution of photochemically processed air masses in the marine boundary layer (MBL). Chebogue Point is approximately 500 km downwind from the eastern sea coast of the United States, and is often impacted by transport of polluted air from the continental source regions.
Northwest Territories, Canada, have been made in the winter to
The correlations between organic nitrates and CO or O3 have been presented by Roberts et al. [1996]. It has been established that CO is a good tracer of continentally derived pollution at this site [Parrish et al., 1993]. This continentally derived pollution contains hydrocarbons (HC) and NO x, the precursors of organic nitrates and ozone, and serves to link the sources of CO with the sources of organic nitrates and ozone. There are distinct differences in the lifetimes of CO, O3 , and organic nitrates, and the factors that they are dependent on. For these reasons, organic nitrates correlate to varying degrees with CO and O3.
Measurements of the organic nitrates; PAN, PPN, methyl-,
ethyl-, 1-propyl-, 2-propyl-, 2-butyl, and 3-pentyl nitrates were made at Chebogue Point, Nova Scotia, during the 1993 NARE campaign. The relationships of these compounds with other measures of continental pollution, for example, CO and 03, are consistent with our view of these compounds as products of HC-NO x photochemistry.
Comparisons of the ratios with that of [2-butyl nitrate]/[butane] were used to test whether the actual data fit the line predicted from rate constants, branching ratios, photolysis rates, and assumed [OH].
Abstract. Measurements of six alkyl nitrates, methyl-, ethyl-, 1-propyl-, 2-propyl-, 2-butyl, and 3-pentyl nitrates, and two peroxycarboxylic nitric anhydrides, peroxyacetic nitric anhydride
(PAN) and peroxypropionic nitric anhydride (PPN) were made
at Chebogue Point, Nova Scotia, during the 1993 North Atlantic
Regional Experiment campaign. The correlation of alkyl nitrates
(RONO2) with carbon monoxide makes it clear that the chemistry at this site is being driven by continental emissions, since RONO2 compounds are known to be by-products of HC-NOx photochemistry.
Comparison of the ratios of alkyl nitrates to their parent hydrocarbons to that of 2-butyl nitrate/butane showed significant
deviations from trends predicted from rate constants, branching
ratios, and loss rates. For ethyl nitrate this deviation is likely due to a combination of many additional pathways  o the intermediate
peroxy radical from the decomposition of larger alkoxy radicals.
The propyl nitrates showed trends that differed from one another; 2-propyl nitrate was a factor of 2 to 3 higher than predicted, while 1-propyl nitrate was a factor of 8 to 10 higher than predicted. The trend of 3-pentyl nitrate/pentane was relatively close (about
50\% lower) to the predicted trend.

\subsubsection*{Biomass burning source}
\citep{Newland2013}
[Alkyl nitrates also have a biomass burning source (Friedli et al., 2001; Simpson et al., 2002). Simpson et al. (2002) report emissions of C1 – C4 alkyl nitrates from savanna fires in Australia, with methyl nitrate being primarily emitted during the ’flaming’ stage and ethyl – butyl nitrates mainly emitted during the ‘smouldering’ stage. They estimate global emissions from this source to be on the order of 8 Gg yr-1]
\citep{Simpson2002}
We report the first observations of the emission of five C1–C4 alkyl nitrates (methyl-, ethyl-, n-propyl-, i-propyl-, and 2-butyl nitrate) from savanna burning. Average alkyl nitrate mixing ratios in the immediate vicinity of three bushfires in Northern Australia were 47–122 times higher than local background mixing ratios. These are the highest alkyl nitrate mixing ratios we have ever detected, with maximum mixing ratios exceeding 3 ppbv for methyl nitrate. Methyl nitrate dominated the alkyl nitrate emissions during the flaming stage of savanna burning, whereas C2–C4 alkyl nitrates were mostly emitted during the smoldering stage. To explain the formation of alkyl nitrates from biomass burning, we propose a reaction mechanism involving the combination of reactive radicals at high temperature. Bearing in mind the uncertainties associated with extrapolating small data sets to much larger scales, alkyl nitrate emissions from global savanna burning are estimated to be on the order of 8 Gg/yr.

\citep{Chuck2002}
Use of the mean sea-to-air fluxes for 0 to 40S (Table 1) and measured atmospheric concentrations results in a lifetime of be- tween 4.5 and 25 days for MeONO2 and between 5 and 10 days for EtONO2 (22). Literature estimates of the atmospheric life- time of MeONO2, based on known atmospheric destruction processes, are poorly con- strained, ranging from 6 to 29 days (23, 24). For EtONO2
, the atmospheric lifetime with respect to photodissociation at the equator and at 40S has been estimated to be 7 and 12 days, respectively, for October (25).

\subsubsection*{Global distribution}
\citep{Newland2013}
[There have been a number of studies of the latitudinal and seasonal variations of alkyl nitrates in marine air (Blake et al., 2003), in the Arctic (Swanson et al., 2003) and in the Antarctic (Beyersdorf et al., 2010).
Blake et al. (2003) report extensive measurements of C1 – C4 alkyl nitrates from the Pacific between 45N and 72S during the PEM-Tropics A and B projects. This study observed an equatorial band between about 8N and 13S in which the oceans are a source of alkyl nitrates. The highest mean methyl and ethyl nitrate measurements were found in this region. The highest mixing ratios of 2-propyl nitrate were found in the most northerly regions of the study (i.e. 40-45N) but it also showed significantly elevated mixing ratios in the same equatorial band as the C1 – C2 nitrates. The high mixing ratios in the northern mid-latitudes are explained by production from anthropogenic polluted outflow. 2-butyl nitrate showed almost no enhancement in equatorial regions and again had the highest mixing ratios in the most northerly regions of the study.
Swanson et al. (2003) and Beyersdorf et al. (2010) observe a strong seasonality of alkyl nitrate mixing ratios in the Arctic and Antarctic respectively. Polar alkyl nitrate mixing ratios show a similar seasonality to the alkanes, with a spring-time peak just before polar sunrise and lows in the summer time. Winter time mixing ratios of 2-propyl and 2-butyl nitrate are about 10 ppt, while methyl and ethyl nitrate are about 5 – 6 ppt. In the summer 2-propyl and 2-butyl nitrate mixing ratios fall close to zero, while methyl and ethyl nitrate mixing ratios fall to about 3 ppt. The amplitude of the seasonal cycle reflects the reaction rate with OH, i.e. it increases with increasing carbon number.]

\subsubsection{Mean AN concentrations}
\begin{table}
\begin{tabular}{cccc}
\hline
Observation & $RONO_2$                                  & Comments & Reference \\
\hline
ITOP?       & mean $\Sigma RONO_2$ 10 ppt               &          & \citep{Reeves2007} \\
???         & mean $\Sigma RONO_2$ 14 ppt               &          & \citep{Roberts1998} \\
remote marine environments & 2-butyl and 2- and 3-pentyl: 0.8-37 pptv & &\citep{Atherton1989} [Atlas, 1988]\\
urban environments         & 2-butyl and 2- and 3-pentyl?: approx. 2 ppb & & \citep{Atherton1989} [Hanst et al., 1982] \\
??                         & MeONO2: 1.2 to 43 pptv     &          & \citep{Chuck2002} \\
??                         & EtONO2: 0.8 to 10 pptv     &          & \citep{Chuck2002} \\
remote marine troposphere  & MeONO2: 0.2 to 50 pptv     & in lit. (2, 8, 9) & \citep{Chuck2002} \\
remote marine troposphere  & EtONO2: 0.3 to 12.2 pptv   & in lit. (2, 8, 9) & \citep{Chuck2002} \\
Pacific Ocean              & MeONO2: 50 pptv            &                   & \citep{Blake2003} \\
airborne                   & $\Sigma RONO_2$: approx 1 ppb         & difference between NOy and (NOx+PAN+HNO3) 
                                                                     is sample closest to the plume center & \citep{Neuman2012} \\
\hline
\end{tabular}
\end{table}

\subsubsection{Comparison methodology}
\citep{Sommariva2008}
[The ratios between the alkyl nitrates and their
main alkane precursors can also be plotted against one another, for example against 2-butylnitrate/ n-butane as in Fig. 4. In this way the modelled and the measured concentrations can be compared without taking into account their variation with time. A similar approach was used in previous studies to investigate the sources of alkyl nitrates and the photochemical processing of polluted air masses (e.g., Bertman et al., 1995; Flocke et al., 1998; Roberts et al., 1998; Stroud et al., 2001; Simpson et al., 2003; Reeves et al., 2007). These studies compared plots of the measured ratios against 2-butylnitrate/n-butane to the ratios calculated with a simple sequential reaction model which assumed a single precursors for each alkyl nitrate. Usually, the underestimation of the ratios for short- chain compounds (propylnitrates/propane, ethylni- trate/ethane) was interpreted as an indication of photochemical sources other than the main alkane precursor.]

\citep{Day2003}
We also show that sumANs are a large part, if not all, of the 'missing NOy' reported in many prior experiments and emphasize that the ratios of sumANs/NOz and of O3/sumANs are especially useful for evaluating chemical models and comparing observations at different sites.

\subsubsection{Modelling studies}
A huge number of modelling studies has been done about alkyl nitrates. In particular, O3 and the RONO2 oxidation products formed from reactions between these NOx and VOC emissions are examined in detail...

A study that enabled us to focus on examination of $O_3$ and $RONO_2$ formation processes has been conducted by \citep{Neuman2012}. After an explosion at the Deepwater Horizon (DWH) offshore oil platform on 20 April 2010 large amounts of oil from the DWH spill reached the ocean surface. Subsequent oil evaporation caused a release of little methane and other light alkanes to the atmosphere forming a $VOC$ mixture atypical for the remote marine environment. At the same the $NO_x$ levels were enhanced in the vicinity of the oil spill due to spill response efforts (ship exhaust, surface oil burning, flaring of recovered gas). Airborne measurements of reactive nitrogen species ($NO_y$ = $NO_x$+$HNO_3$+$PAN$(peroxyacetul nitrate at the bottom)+$RONO_2$) and $VOCs$ in the marine boundary layer during daytime downwind from the DWH area showed that oxidation of alkanes with between 5 and 10 carbons per molecule ($C_5$–$C_{10}$) dominated $O_3$ formation chemistry in the plume. This circumstance led to a rapid and nearly complete oxidation of $NO_x$ to $RONO_2$ with mostly 6-10 carbons per molecule ($C_6$–$C_{10}$) effectively terminating $O_3$ production. On 10 June, $O_3$ enhancements in the DWH plumes were less than 15 ppbv, even after $NO_x$ was completely oxidized.

$O_3$ production efficiencies in the DWH plume were 6 plus minus 0.5 ppbv $O_3$ per ppbv of $NO_x$ oxidized after 1 h of transport. These values are comparable with those measured in U.S. urban areas [Trainer et al., 1995; Neuman et al., 2009], but smaller than those measured in the petrochemical plumes with $OH$ loss rates dominated by alkenes (10-18 ppbv $O_3$ per ppbv of $NO_x$) [Ryerson et al., 2003].

Temporal changes in reactive nitrogen partitioning in the DWH plume (Figure \ref{fig:Neuman2012}) nicely illustrate that in fresh part of the plume (Figure ??? a) $NO_x$ accounted for over 80\% of plume $NO_y$, in aged plumes (Figure ??? b), $NO_x$ oxidation was nearly complete, but NOx, PAN, HNO3, and the sum of eight $C_5$–$C_{10}$ $RONO_2$ accounted for only 22\% of plume $NO_y$. It means that $C_6$–$C_{10}$ $RONO_2$ account for 78\% of the aged plume $NO_y$\citep{Neuman2012}. This result differs dramatically from previous studies, where the estimated contribution of $RONO_2$ to $NO_y$ was much lower: approximately 4\% in urban and petrochemical plumes [Neuman et al., 2002; Ryerson et al., 2003], approx. 10\% in rural areas [Parrish et al., 1993], a few percent in urban air masses [O’Brien et al., 1997; Flocke et al.,1998; Schneider et al., 1998; Talbot et al., 2003; Ryerson et al., 2003], and up to 10–20\% of NOy in some studies [Day et al., 2003; Perring et al., 2010].

\begin{figure}[h!]\label{fig:Neuman2012}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neuman2012_Fig.4}.png}
\caption{Time series of reactive nitrogen species measured on 10 June 2010 in the MBL in plumes (a) 10 km and (b) 50 km downwind from DWH. $NO_x$, $NO_y$, and $HNO_3$ were measured once per s and $PAN$ was measured once per 2 s. $C_1$–$C_5$ alkyl nitrates were measured in whole air samples at locations indicated by the circles. The sum of alkyl nitrates (black lines) is inferred from the difference between $NO_y$ and ($NO_x$+$HNO_3$+$PAN$) \citep{Neuman2012}.}
\end{figure}

The first 3-D chemical transport  model simulations examining the impact of methyl and ethyl nitrates on tropospheric ozone chemistry in remote regions have been conducted by \citep{Neu2008}. In their study, they incorporated methyl and ethyl nitrates in the chemical mechanism of the UCI CTM [Carver et al., 1997] and by matching their observed atmospheric abundances derived a total oceanic emission flux of these two nitrates (0.35 Tg of N per year). Comparing two simulations with and without spatially uniform emissions of methyl and ethyl nitrates they found that the largest impact of oceanic alkyl nitrates is seen over the equatorial Western Pacific (Figure ???). Atmospheric circulation (convergence of the trade winds) in this region isolates the marine boundary layer, resulting in sustained photochemical loss of $O_3$ (Figure ??? c) and at the same time accumulation of alkyl nitrates (Figure ??? a), which drive up boundary layer $NO_x$ abundances by as much as 250\% (Figure ??? b). Such combination of factors leads to increases in $O_3$ of up to 20\% (Figures ??? c and d), with the largest increases occurring from May to October.

\begin{figure}\label{fig:Neu2008}
\caption{1}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.2_MeONO2_OCT_colorbar}.png}
\end{figure}
\begin{figure}
\caption{1}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.2_MeONO2_OCT}.png}
\end{figure}
\begin{figure}
\caption{1}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.3_OCT_NOxdiff}.png}
\end{figure}
\begin{figure}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.3_OCT_O3contol_O3diff}.png}
\caption{October monthly mean mixing ratios of (a) methyl nitrate ($MeONO_2$, $CH_3ONO_2$) and (c) $O_3$, and percent difference (perturbation - control) in (b) $NO_x$, (d) $O_3$. All averaged over the boundary layer (below 850 mb). Adapted from \citep{Neu2008}}
\end{figure}

\citep{Atherton1989}
The transport from urban areas can lead to alkyl nitrate concentrations of 4-83 pptv after 120- 240 hours. These levels are far higher than those possible from marine in-situ production. Additionally, alkyl nitrates may be more abundant than PAN in remote marine environments.
Model: box
Chemistry: explicit chemical reaction mechanism
[Carter et al., 1986. A variety of alkyl nitrates, including methyl-, butyl-, pentyl-, hexyl-, heptyl-, octyl-, and nonyl-nitrates are allowed to form from fourteen different alkyl peroxy radicals. + ocean emission sources fluxes of c
for ethene, ethane, propene, and propane, NO, NO2
Initial conditions: clear sky conditions, a quantum yield of 1, solar zenith angle of 45 degrees, and total ozone column of 8*10 cm-2. Daytime+nigthtime, invertion height, dilution. 298 K. No deposition.
Scenarios: Two, MLOW and MHIGH, represent the marine in-situ production of alkyl nitrates. The total hydrocarbon concentration was 17.6 ppbC. MLOW had NO = 4 pptv and NO2 = 4.6 pptv based on measurements over the Pacific Ocean. MHIGH had NO = 46 pptv, NO2 = 54 pptv, and PAN = 50 pptv. The two other scenarios, UHIGH and ULOW, simulate the transport of species from urban to remote environments. UHIGH contained 2150 ppbC and 140 ppbv NOx, and ULOW contained 270 ppbC and 18 ppbv NOx, giving each a HC/NOx ratio of 15.
Simulation time: 12 hours MLOW and MHIGH, 240 hours UHIGH and ULOW.
Results:
- The total alkyl nitrate concentration present after 12 hours is less than 0.4 pptv for MHIGH and less than 0.1 pptv for MLOW. These levels are much lower than observations, indicating that marine in-situ production can account for only a fraction of the total marine alkyl nitrate concentration.
- IHIGH The total alkyl nitrate concentration peaks during the first 6 hours and then drops to 3.1 ppbv after 12 hours. This value is com- parable with the 2 ppbv alkyl nitrate concentration de- termined in Los Angeles [Hanst et al., 1982]
- The model predicts alkyl nitrate concentrations of 21-83 pptv for UHIGH and 4-12 pptv for ULOW for the hours 120-240.
- General: The model predictions of 12-83 pptv (af- ter 120 hours) and 4-21 pptv (after 240 hours) are of the same order as observations.

\citep{Sommariva2008}
Validation: typical urban plume in the North-East of the United States, measurements taken during three flights of the NOAA WP-3D aircraft, which sampled a plume from the New York City area during the New England Air Quality Study (NEAQS) 2004 campaign
- Seven alkyl nitrates (ethyl- nitrate, i-propylnitrate, n-propylnitrate, 2-butylni- trate, 2-pentylnitrate, 3-pentylnitrate, 3-methyl-2- butylnitrate) were measured during these flights together with a full suite of VOCs and other chemical parameters, such as CO, O3,NOx, HNO3 and PANs (Neuman et al., 2006).
- The model results have been qualitatively compared to measurements of alkyl nitrates taken on 20–22 July 2004, when the NOAA WP-3D aircraft made three flights aimed to inter- cept the New York City plume.
Model: subset! of MCM v3.1 (The MCM is one of the most detailed chemical mechanism for tropospheric chemistry available)
- to study the formation processes and photochemical evolution of alkyl nitrates
Chemistry: oxidation mechanisms of 64 VOCs (50 hydrocar-
bons and 14 oxygenates) plus CH4 and CO, The inorganic chemistry mechanism was taken from the IUPAC recommendation (Atkinson et al., 2003). A dimethyl sulphide (DMS) mechanism (Sommariva et al., 2004) was also included. The photolysis rates were calculated using the MCM parametrization (Saunders et al., 2003) for end of July 2004 at the latitude and longitude of Boston, MA. only gas-phase!!! Measurements made during NEAQS 2004 showed that there was very little removal of, for example,HNO3 by dry deposition in plumes decoupled from the marine boundary layer during transport from New York City (NYC) to the North Atlantic (Neuman et al., 2006). Therefore, removal by dry deposition was neglected in the model.
Initialisation: measurements taken in the Boston metropolitan area during the 2002 and 2004 NEAQS campaigns. The initial concentrations of VOCs were determined using emission ratios; CH4 and O3 were set to typical values for urban conditions (1800 and 90 ppb) (Table 1). The model has been initialized using measurements and VOCs emission ratios from the 2002 and 2004 New England Air Quality Studies.
Simulation time: 5 days.
Scenarios: 2: A (higher OH, good with obs, reference), B
Time: 20-22 July.
Results:
- While long-chain (C5) alkyl nitrates are produced for 90\% or more from the oxidation of a single parent alkane, short-chain <C4 alkyl nitrates can be formed from several precursors. The relative importance of each production route has been quantitatively determined thanks to the high level of chemical detail provided by the MCM. These secondary routes
to the formation of alkyl nitrates include the oxidation of longer-chain alkanes and oxygenated intermediates, like carbonyls, peroxides and carboxylic acids.
- The best agreement is in the case of 3-methyl-2-butylnitrate and the discrepancy between the modelled and the measured ratios clearly increases with decreasing carbon number .
- RH/AN, Except in the case of ethylnitrate, the model can reproduce the measured rate of increase in the alkyl nitrate to alkane ratios fairly well, which suggests that the kinetic parameters used in the MCMmodel are overall correct.
- By considering together Figs. 3 and 4 it can be reasonably concluded that the chemistry of alkyl nitrates is well represented in the MCM model.
- Using the explicit description of the chemistry provided by the MCM, it has been possible to determine quantitatively the photochemical formation and the evolution of alkyl nitrates and the relative importance of their different formation pathways. The results are summarized in Table 2. Large alkyl nitrates are produced almost exclusively from the reaction of a single parent
alkane. Smaller <=C4 alkyl nitrates also have other significant precursors, like longer-chain alkanes and alkenes. For n-propylnitrate and ethylnitrate, pro- panal and butanal, which are largely emitted, are important sources within the first 24 h. For all alkyl nitrates, the role of organic peroxides and car- boxylic acids formed by the reactions of peroxy radicals with HO2 at low [NOx] increases with time. These species contribute to sustain the production of alkyl nitrates far from the emission points of the parent VOCs. The relative importance of the different pathways
and how they change with time are dependent on the initial conditions of the model. In a scenario with lower O3 and CO and higher NOx the
importance of the direct routes (alkane ?OH) is typically lower during the first two–three days, and the contributions of peroxides and carboxylic acids become significant later and are on average less important. These results suggest that secondary chemistry in the formation of the alkyl nitrates, particularly oC4, cannot be neglected and should be included in the chemical models, due to its potential impact on ozone-controlling strategies.

\citep{Newland2013}
Model1: 2-D atmospheric model
- to study the formation, loss and transport of the alkyl nitrates to the Arctic
- to derive northern hemisphere emissions of the six alkanes.
Chemistry:
- (with fixed NOx and OH field taken from full chemistry run with model by Hough and Johnson (1990))
Validation of ANs: comparison with measurements of alkyl nitrates in the Arctic.
Performance: recreate both the measured seasonality of the alkyl nitrate, and the peak mixing ratio, reasonably well.
Simulation time: 6 days?
Model2: box
Chemistry:
- alkanesmethane, ethane, propane, butane and pentane and ANs explicitly described (RO+NO2 too)
Model3: firn model
- to derive atmospheric histories of six alkanes and six alkyl nitrates from measurements in firn air from the NEEM site, Greenland (methyl, ethyl, 2-propyl, 2-butyl, 2+3-pentyl and 3- methyl-2-butyl nitrate), 1950-2008.
Results:
- There are two ways in which NOx mixing ratios can limit alkyl nitrate production: either through air masses being NO limited with respect to the formation of a given alkyl nitrate or by the NOx in an air mass being 'used up' (track back to Reeves)
-- Alkyl nitrate production becomes NOx saturated above a given NO mixing ratio. This limit increases with increasing carbon number.
-- The time taken for an air mass to 'use up' all the initial NOx is dependent mainly on the initial NOx mixing ratio.
- At high [NO2] (> 27 ppb), methyl nitrate production is dominated by the reaction of the alkoxy radical with NO2
- The regions that contribute most to modeled alkane and alkyl nitrate mixing ratios in the Arctic are Europe and Russia
- The importance of the RO + NO2 pathway compared to RO2 + NO is entirely dependent on the branching ratio (alpha3) of the RO2 + NO reaction pathway. The smaller alpha3 is, the more important the RO + NO2 pathway is.
-- At 10 ppb of NO2 the RO + NO2 pathway accounts for 30 – 60\% of methyl nitrate production. At 100 ppb it accounts for 85 – 95\% of methyl nitrate production and almost 20\% of ethyl nitrate production.
- The rate of alkyl nitrate production in a polluted air mass is proportional to the OH concentration in that air mass.

\subsubsection*{Other}

To sum up, alkyl nitrates are produced as a minor branch of the reaction of peroxy radicals with NO \citep{Day2003}. Since these compounds are removed from the troposphere by photolysis and OH oxidation they can provide information on photochemical evolution of an air mass through relationships with their parent hydrocarbons \citep{Bertman1995}. Acting as reservoir species for reactive nitrogen (NOy) and having a relatively long atmospheric lifetimes alkyl nitrates play an important role in long range transport of NOx. Furthermore, alkyl nitrates link to ozone and carbonyl production making them useful tracers for both species \citep{Worton2010}. Despite their relative importance alkyl nitrates are usually not included in the climate-chemistry models, for example in UKCA (Reeves, personal communication), which potentially could lead to discrepancies between models and observations described before.

[Source characterization and photochemical processes in
transported air masses have been analyzed by examining
RONO2 formation from anthropogenic emissions of alkane
precursors [Simpson et al., 2003; Madronich, 2006; Simpson
et al., 2006; Reeves et al., 2007; Worton et al., 2010]. (Neuman2012)]

They are produced photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990} or emitted directly from equatorial oceans \citep{Blake2003} and biomass burning \citep{Simpson2002}. Being involved in the ozone production chemistry, these species can act as reservoirs of nitrogen oxides which availability largely affects the production of ozone \citep{Reeves2007}. Since the alkyl nitrates have atmospheric lifetimes ranging from several days to about a month, they can redistribute nitrogen oxides presented over polluted continental regions to clean remote marine environments, and therefore are considered as one of the tracers for determining the anthropogenic influence \citep{Atherton1989, Reeves2007, Worton2005}.

Owing to the existence of the relationship between alkyl nitrates and their parent hydrocarbons, measurements of their concentrations can provide information on the extent of photochemical processing in the air mass. \cite{Bertman1995} developed a kinetic approach to describe this relationship through the evolution of the ratio of alkyl nitrate to its parent hydrocarbon as a function of time. This approach has been extensively used to study air mass photochemical ageing at several sites, in North America \citep{Bertman1995, Roberts1998}, China \citep{Simpson2006}, the British Isles \citep{Worton2010} and over the Pacific \citep{Simpson2003} and North Atlantic Ocean \citep{Reeves2007}.

\cite{Bertman1995} found that the theoretical ratios of  2-butyl nitrate to n-butane and 2- and 3-pentyl nitrates to pentane strongly agree with those measured from the ambient air showing no evidence for primary sources of these nitrates. On the contrary, measurements of the other ratios, ethyl/ethane, n-propyl nitrate/propane, and 2-propyl nitrate/propane, appeared to be significantly higher than predicted, relative to the 2-butyl nitrate/n-butane ratio, meaning that there should be an additional source of these alkyl nitrates. \cite{Roberts1998} generally approved the theory by comparing Bertman's conclusions with the results from a new observational dataset.

Using methyl nitrate as a tracer of marine $RONO_2$ \cite{Simpson2006} revealed that the South China Sea in not a region of strong $RONO_2$ emissions suggesting that photochemical rather marine production of alkyl nitrates is dominant at the site.

\cite{Worton2010} found that the photochemical source of the short-chain alkyl nitrates originates mostly from the photochemical oxidation and decomposition of longer-chain compounds rather than from the oxidation of the parent hydrocarbons as it is thought in the previous studies.

The brief description of the measurement sites and data is presented in section \ref{sec:data}. In section \ref{sec:method} the methodology used in this work is defined. Section \ref{sec:res} contains the results. Section \ref{sec:discuss} discusses the results, particularly focusing attention on the analysis of the discrepancy between the observational data and the results of theoretical modelling. Conclusions constitute section \ref{sec:conclusion}.

\section{Methodology} \label{sec:method}
\subsection{Model development and description}
A zero-dimensional (box) model is developed to investigate the impact of alkyl nitrate chemistry on ozone production efficiency. Its code is written using FACSIMILE for Windows. FACSIMILE is a special high-level programming language which is a powerful modelling tool designed to efficiently solve differential equations including those resulting from a set of chemical reactions.

! \citep{Newland2013}
[The model is written using the programming language FACSIMILE (Curtis and Sweetenham, 1987) and the differential equations are assembled and integrated using the Gear’s method (1971). The mixing ratios of the input species in the model are driven entirely by the emissions, the loss processes and the transport with no redistribution of the burden to ensure that mixing ratios agree with observations.]

The model is intended to represent the steady state of the troposphere under mean summer daytime meteorology and clear skies. To achieve that a number of assumptions have been made. First, the model assumes a uniform mixing of individual constituents of the troposphere, second, there are no atmospheric transport (no 'winds') and no emissions of species into the box meaning that the time evolution of each species is controlled by the chemical interactions alone and determined by their initial concentrations. The air temperature and relative humidity are assumed to be $25^{\circ}C$ and 70\% respectively and are fixed in all model runs. Other assumptions concerning the chemical mechanism and photolysis rates can be found in the following sections.

\subsection{Chemical mechanism}\label{sec:chem_mech}
The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, retrieved via website: http://mcm.leeds.ac.uk/MCM. On a whole it includes 90 species and 165 reactions. The mechanism is developed to represent the chemistry that governs the production and loss of ozone, hydrocarbons and alkyl nitrates. These include:

! \citep{Newland2013}
[The majority of the kinetic reaction rates and photochemical data used in the full chemistry model were taken from the IUPAC Subcommittee for Gas Kinetic Data Evaluation (http://www.iupac-kinetic.ch.cam.ac.uk; Atkinson et al., 2004; Atkinson et al., 2006; Atkinson et al., 2008). If the required data were not provided by IUPAC then data were taken from the NASA / JPL ‘Chemical Kinetics and Photochemical Data for Use in Atmospheric Studies: Evaluation No. 17’ (Sander et al., 2011). Both these sources consider the full range of literature available on each reaction and provide a recommendation based on a synthesis of this data. There are sometimes minor differences between the recommendations of each organisation.]

\begin{itemize}
\item the formation of ozone as a result of the photolysis of $NO_2$ followed by the reaction of ground-state oxygen atom ($O(^3P)$) with molecular oxygen and a third body, M ($N_2$ or $O_2$);
\item the loss of ozone through its photolysis and reactions with $O(^3P)$, $OH$, $HO_2$ and $NO$, among which the last is of particular interest because it gives back $NO_2$;
\item the formation of alkylperoxy radicals ($RO_2$) from oxidation of the hydrocarbons;
\item the competition of $NO$ and $HO_2$ for reaction with $HO_2$ and $RO_2$;
\item the formation of alkyl nitrates which adds up to the competition of $NO$ and $HO_2$ for reaction with the peroxy radicals;
\item the loss of alkyl nitrates through their photolysis and reaction with $OH$;
\item the subsequent reactions of hydroperoxides and aldehydes formed from the organic reactions mentioned above.
\end{itemize}
The full list of reactions can be found in Appendix ?? \ref{sec:appendix1}. It is important to mention that due to purely theoretical purpose of this research and its focus on investigation of the impact of alkyl nitrate chemistry on ozone production some of the reactions that lead to formation of species that serve as a sink for $NO_x$ are not included into the chemical mechanism. However, these reactions play an important role in the chemistry of the troposphere, and below here is a description of their role and anticipated effect of their exclusion.
\begin{itemize}
\item The formation of the nitric acid ($HNO_3$) is a major loss process for $NO_x$ during daytime in the atmosphere. Being a 'sticky' molecule $HNO_3$ readily adsorbs to surfaces, particularly if there is water on the surface, and deposits either in precipitation (wet deposition) or in dry form (dry deposition). Because of the high deposition velocity, dry deposition of $HNO_3$ can be responsible for much of the removal of inorganic nitrogen from the troposphere \citep{Finlayson-Pitts2000}, which means that exclusion of $HNO_3$ (formation and removal)/(chemistry) from the chemical mechanism potentially leads to overproduction of ozone in the model, especially at high $NO_2$ concentrations because the competition between $NO_2$ photolysis and its reaction with $OH$ radical ($NO_2 + OH \xrightarrow{M} HNO_3$) is ignored.
\item Since this research is focused on the daytime chemistry some reactions that are believed to be significant in nighttime are not included into the chemical mechanism of the model. Among them are reactions that involve the nitrate radical ($NO_3$) and dinitrogen pentoxide ($N_2O_5$).
\begin{itemize}
\item $NO_3$ radical rapidly photolyzes during the daylight hours, but in nighttime its concentration can increase to measurable levels \citep{Atkinson2008}. This increase in concentration and a lesser availability of the OH radical (which sources are mostly photolytic in nature) make $NO_3$ radical a major contributor to the night chemistry of organics in the troposphere \citep{Finlayson-Pitts2000}. However, exclusion of $NO_3$ radical reactions from our chemical mechanism should not have a huge impact on the results anyway since the mechanism considers mainly decomposition of alkanes with which $NO_3$ radical reacts to a generally minor extent \citep{Atkinson2008}???.
\item ! the nighttime NO3 radical reaction is of minor significance in term of the overall removal of the alkanes (Atkinson, 1990a)
\item In addition to reacting with organics, $NO_3$ also reacts with $NO_2$, forming dinitrogen pentoxide ($NO_3 + NO_2 \xleftrightarrow{M} N_2O_5$) which in turn is an important nighttime source of nitric acid through its rapid hydrolysis on wet surfaces and aerosol particles ($N_2O_5 + H_2O \xrightarrow{surf., aerosol} 2HNO_3$) \citep{Finlayson-Pitts2000}. Exclusion of $N_2O_5$ from the chemical mechanism assumes no cycling of nitrogen oxides within the troposphere through $NO_3$ and $N_2O_5$, eventually forming $HNO_3$, which are not included into the mechanism anyway.
\item ! In any case, because of this equilibrium, sink for N2O5 such as hydrolysis are, in essence, also sinks for NO3 as well.
\end{itemize}
\item Another species involved in $NO_x$ tropospheric chemistry is the nitrous acid ($HONO$). The major source of nitrous acid is believed to be heterogeneous 'dark' reactions of $NO_2$, including hydrolysis (reaction with water vapour) of $NO_2$ on aerosol and particulate matter surfaces in nighttime. During the daytime, $HONO$ can be formed by the reaction of $OH$ radical with $NO$ ($OH + NO \xrightarrow{M} HONO$). However, since $HONO$ strongly absorbs UV radiation, it rapidly photolyzes, and under particular conditions (when the nighttime $HONO$ concentration reaches almost 15 ppb \citep{Finlayson-Pitts2000}), photolysis of nighttime-generated $HONO$ can become a major source of $OH$ radicals in the early morning hours ($HONO + hv \rightarrow OH + NO$) \citep{Lammel1996}. Therefore exclusion of $HONO$ from the chemical mechanism potentially leads to underestimation of $OH$ concentrations, but having neither formation nor removal of $HONO$ should balance this effect out???.
\item Peroxyacetyl nitrate ($CH_3C(O)OONO_2$, PAN) is the most abundant organic nitrate in the troposphere \citep{Hewitt1994}. A classic precursor to the PAN is acetaldehyde ($CH_3CHO$) which oxidation produces PAN through the following reaction sequence:
\begin{itemize}
\item[] $CH_3CHO + OH \rightarrow CH_3CO + H_2O$
\item[] $CH_3CO + O_2 \xrightarrow{M} CH_3C(O)OO$
\item[] $CH_3C(O)OO + NO_2 \xrightarrow{M} CH_3C(O)OONO_2$.
\end{itemize}
Of all the possible fates for PAN in the atmosphere, thermal decomposition is usually the most important because the rate constant for PAN decomposition strongly depends on temperature. At low temperatures it is quite stable, but at higher temperatures it decomposes giving $NO_2$ and the acetylperoxy radical ($CH_3C(O)OO$) 
\begin{itemize}
\item[] $CH_3C(O)OONO_2 \xleftrightarrow{M} NO_2 + CH_3C(O)OO$.
\end{itemize}
This temperature dependence has important implications for the role of PAN. When it is formed at lower temperatures (e.g., at high altitude) or transported into colder regions, PAN can act as an $NO_x$ reservoir. Observations of PAN and the sum of reactive nitrogen ($NO_y$) confirm it revealing that PAN can constitute up to 90\% of the total $NO_y$ budget at higher northern latitudes or higher altitudes (e.g., \citep{Jacobi2000}). When an air mass containing PAN is transported into warmer regions, $NO_2$ tied up in PAN can be released back and cause the formation of $O_3$ and $HNO_3$ far away from the original $NO_x$ source. Moreover, if sufficient $NO$ is present, $CH_3C(O)OO$ released from PAN can react in the following manner:
\begin{itemize}
\item[] $CH_3C(O)OO + NO \rightarrow CH_3C(O)O + NO_2$
\item[] $CH_3C(O)O \rightarrow CH_3 + CO_2$
\end{itemize}
followed by the reaction of $CH_3$ with $O_2$, etc. to form $HCHO$ and $HO_2$, or $HCHO$, $CH_3OH$ and $CH_3OOH$ under low-$NO_x$ conditions. As a result of this chemistry, PAN can act as an accelerator for photochemical smog formation \citep{Finlayson-Pitts2000}. Exclusion of PAN from the chemical mechanism of the model means that we omit the global transport of $NO_x$ provided by PAN, and therefore do not fully estimate tropospheric ozone production. However, since the focus of this research is on the impact of alkyl nitrate chemistry on ozone such a miscount is acceptable due to purely theoretical purpose of the research.
\item Inclusion of all chemistry listed above into the model is a subject for future work.
\end{itemize}

The constructed chemical mechanism has a limited representation of organic chemistry. It includes oxidation of only seven alkanes: methane ($CH_4$), ethane ($C_2H_6$), propane ($C_3H_8$), n-butane ($nC_4H_{10}$), i-butane ($iC_4H_{10}$), n-pentane ($nC_5H_{12}$) and i-pentane ($iC_5H_{12}$). $CH_4$ undergoes a full oxidation in the model, but oxidation of other six alkanes follows the reaction sequence:
\begin{itemize}
\item[] $RH + OH \rightarrow RO_2 + H_2O$
\item[] $RO_2 + HO_2 \rightarrow ROOH + O2$
\item[] $RO_2 + NO \rightarrow RO + NO_2$
\item[] $ROOH + hv \rightarrow RO + OH$
\item[] $ROOH + OH \rightarrow products$
\item[] $RO + O_2 \rightarrow products$
\end{itemize}
and then is 'terminated' after formation of certain species. Ethane oxidation is terminated after formation of the acetylperoxyl radicals ($C_2H_3CO_3$), propane - after formation of the propyldioxy radicals ($C_3H_5O_3$), n-butane - after formation of ethanal ($CH_3CHO$), butanal ($C_3H_7CHO$), methyl ethyl ketone ($C_2H_5COCH_3$), and 1-($\lambda^1$-oxidanylmethoxy)-2-methoxy-ethane ($C_4H_9O_3$), i-butane - after formation of 2-methylpropanal ($iso-C_3H_7CHO$) and acetone ($CH_3COCH_3$), n-pentane - after formation of pentanal ($C_4H_9CHO$), methyl propyl ketone ($n-C_3H_7COCH_3$), and diethyl ketone ($(C_2H_5)_2CO$), i-pentane - after formation of 2-methylbutanal ($sec-C_4H_9CHO$) and isopropyl methyl ketone ($iso-C_3H_7COCH_3$). It has been done to limit the number of reactions in the mechanism, which otherwise would include 1678 reactions describing evolution of 539 species, and therefore would be computationally too expensive. For the same reason all self-reactions of organic peroxy radicals ($RO_2$) are omitted as well. To avoid a huge impact of incomplete oxidation of alkanes on the results the model's simulation time has been restricted to 24 hours.

\citep{Newland2013}
[The alkanes and alkyl nitrates studied in this work were all explicitly described in the model with oxidation of the alkane giving a peroxy radical and reaction of the peroxy radical with NO giving an alkyl nitrate.]

Lastly, the chemical mechanism of the model does not include halogen chemistry. Despite that halogens (chlorine ($Cl$), bromine ($Br$) and iodine ($I$)) are very reactive chemical compounds mostly known by their role in anthropogenic stratospheric ozone-depletion chemistry, and that they are directly (for example, reaction of the $Cl$ radical with hydrocarbons (e.g., $CH_4$)) and indirectly (coupling with the cycles of sulfur and nitrogen that eventually influence $O_3$, $OH$ and $NO_x$ concentrations) involved in the tropospheric chemistry \citep{VonGlasow2014,Platt2003}, they are excluded from the chemical mechanism to keep it more transparent.
\subsection{Photolysis}
To estimate photolysis rates for a given geographical location, one must take into account the latitude and season, as well as the time of the day. For the purpose of this research 24-hour average photolysis rates are calculated externally (outside FACSIMILE program) for 1 July using the formula and parametrisations from the MCM v3.3. The formula that describes variation of photolysis rate with solar zenith angle for the core photolysis reactions in the MCM was obtained by its authors using cross-section and quantum yield data and two stream isotopic scattering model described in \citep{Hayman1997}. Their calculations were performed for clear sky conditions at an altitude of 0.5 km for 1 July at a latitude of 45$^{\circ}$N, resulting in the following formula:
\begin{equation} \label{eq:MCMphotolysis}
\renewcommand{\theequation}{\arabic{equation}}
J = l(cos\chi)^mexp(-nsec\chi)
\end{equation}
where $J$ - photolysis rate coefficient, $\chi$ - solar zenith angle, $l$, $m$ and $n$ - optimising parameters \citep{Jenkin1997,Saunders2003}.

The values of solar zenith angle are calculated using NOAA Solar Calculator (http://www.esrl.noaa.gov/gmd/grad/solcalc/).

\subsection{Design of experiments}
To investigate the influence of alkyl nitrate chemistry on (potential changes in) tropospheric $O_3$ production, box model runs were performed at a series of different $NO_x$ and $VOCs$ initial concentrations, so that an $O_3$ isopleth plot could be constructed, similar to those found in \citep{Dodge1977} and \citep{Sillman1999}.   Table \ref{tab:setupO3CO} shows initial concentrations of 'independent'/global parameters and indicates which of them are fixed or allowed to vary during the model run. Table \ref{tab:setupNOx} and Table \ref{tab:setupVOCs} show the initial concentrations of $NO_x$ and $VOCs$ respectively that are used at a series of experiments.

The lower limit of the range of $NO_x$ and $VOCs$ initial concentrations is based on observational data from Mauna Loa station (Hawaii, US) which is believed to be representative for conditions of remote troposphere. The upper limit is based on data from 
Marylebone Road station (London, UK) which is a kerbside monitoring site located in central London, in close proximity to a 6 lane road, with very high traffic flow and frequent congestion \citep{VonSchneidemesser2010}, therefore being representative for conditions of urban troposphere.

Two experiments have been performed. The aim of the first one is to reveal the total impact of alkyl nitrate chemistry on $O_3$ production, the aim of the second one is to estimate the contribution of individual alkyl nitrates to changes in $O_3$ production (sensitivity!).

For the first experiment the model was run under two configurations, (1) with all alkyl nitrate chemistry switched on, and (2) with all alkyl nitrate chemistry switched off. In every model run $VOCs$ concentrations were (changed according to the data from Table \ref{tab:setupVOCs} and) were fixed throughout the run, while the initial concentrations of $NO_x$ were (changed according to the data from Table \ref{tab:setupNOx} and were) allowed to vary throughout the run. As a result, the first experiment consists of 253 model runs including 11 runs (varying $NO_x$) with inorganic chemistry only.

For the second experiment the model was run using the same approach concerning variations in $NO_x$ and $VOCs$ concentrations, but changes in alkyl nitrate chemistry were treated differently. In every model run reactions of individual alkyl nitrates or their groups (defined in Table \ref{tab:RHandANS}) were either switched on or switched off, resulting in other 847 additional model runs.

!! C2H5NO3 ethylnitrate, IC3H7NO3 i-propylnitrate, NC3H7NO3  n-propylnitrate, SC4H9NO3 2-butylnitrate, PEBNO3 2-pentylnitrate;PECNO3 3-pentylnitrate, IPEBNO3 3-methyl-2-butylnitrate

\begin{table} % RH and ANs
\caption{Alkyl nitrates and their parent alkanes}\label{tab:RHandANS}
\centering
\begin{tabular}{llll}
\hline
Parent alkane& Structural formula          & Name                     & Group\\
\hline
$CH_4$       & $CH_3ONO_2$                 & methyl nitrate           & $CH_3ONO_2$ \\
\hline
$C_2H_6$     & $CH_3CH_2ONO_2$             & ethyl nitrate            & $C_2H_5ONO_2$ \\
\hline
$C_3H_8$     & $(CH_3)_2CHONO_2$           & iso-propyl nitrate       & $\Sigma C_3H_7ONO_2$\\
             & $CH_3(CH_2)_2ONO_2$         & n-propyl nitrate         & \\
\hline
$nC_4H_{10}$ & $CH_3(CH_2)_3ONO_2$         & n-butyl nitrate          & $\Sigma nC_4H_9ONO_2$\\
             & $CH_3CH_2CH(CH_3)ONO_2$     & sec-butyl nitrate        & \\
\hline
$iC_4H_{10}$ & $CH_3CH(CH_3)CH_2ONO_2$     & iso-butyl nitrate        & $\Sigma iC_4H_9ONO_2$\\
             & $CH_2C(CH_3)CH_2ONO_2$      & tert-butyl nitrate       & \\
\hline
$nC_5H_{12}$ & $CH_3(CH_2)_4ONO_2$         & n-pentyl nitrate         & $\Sigma nC_5H_{11}ONO_2$\\
             & $CH_3(CH_2)_2CH(CH_3)ONO_2$ & 2-pentyl nitrate         & \\
             & $???$                       & 3-pentyl nitrate         & \\
\hline
$iC_5H_{12}$ & $???$                       & 2-methylbutyl nitrate    & $\Sigma iC_5H_{11}ONO_2$\\
             & $(CH_3)_2CH(CH_2)_2ONO_2$   & 3-methyl-2-butyl nitrate & \\
             & $(CH_3)_2C(CH_2CH_3)ONO_2$  & 2-methyl-2-butyl nitrate & \\
\hline
\end{tabular}
\end{table}

!! For long-lived species, averaging may be reasonable.

\begin{table} % model setup: O3, CO
\caption{Fixed and variable 'independent' parameters used in box model runs}\label{tab:setupO3CO}
\centering
\begin{tabular}{ccc}
\hline
Parameter & Initial mixing ratio & Type \\
\hline
$H_2O$    & 2\%                  & Fixed    \\
$O_2$     & 21\%                 & Fixed    \\
$N_2$     & 78\%                 & Fixed    \\
$H_2$     & 525 ppb              & Variable \\
$O_3$     & 40 ppb               & Variable \\
$CO$      & 100 ppb              & Variable \\
\hline
\end{tabular}
\end{table}

\begin{table} % model setup: NOx
\caption{Model setup: NOx initial concentrations (ppb)}\label{tab:setupNOx}
\centering
\begin{tabular}{ccc}
\hline
$NO$      & $NO_2$      & NOx  ($=NO+NO_2$) \\
\hline
0.0015    & 0.0035      & 0.005 \\
0.007     & 0.018       & 0.025 \\
0.015     & 0.035       & 0.05  \\
0.03      & 0.07        & 0.1   \\
0.07      & 0.18        & 0.25  \\
0.15      & 0.35        & 0.5   \\
0.25      & 0.5         & 0.75  \\
0.3       & 0.7         & 1     \\
0.7       & 1.8         & 2.5   \\
1.5       & 3.5         & 5     \\
3         & 7           & 10    \\
\hline
\end{tabular}
\end{table}

\begin{table} % model setup: VOCs
\caption{Model setup: VOCs concentrations (ppb)}\label{tab:setupVOCs}
\centering
\begin{tabular}{ccccccccc}
\hline
Experiment number & $CH_4$ & $C_2H_6$ & $C_3H_8$ & $nC_4H_{10}$ & $iC_4H_{10}$ & $nC_5H_{12}$ & $iC_5H_{12}$ \\
\hline
A  & 1780 &	0.620 &	0.100 &	0.050 &	0.250 &	0.250 &	0.250 \\
B  & 1807 &	1.258 &	0.390 &	0.245 &	0.325 &	0.275 &	0.425 \\
C  & 1834 &	1.896 &	0.680 &	0.440 &	0.400 &	0.300 &	0.600 \\
D  & 1861 &	2.534 &	0.970 &	0.635 &	0.475 &	0.325 &	0.775 \\
E  & 1888 &	3.172 &	1.260 &	0.830 &	0.550 &	0.350 &	0.950 \\
F  & 1915 &	3.810 &	1.550 &	1.025 &	0.625 &	0.375 &	1.125 \\
G  & 1942 &	4.448 &	1.840 &	1.220 &	0.700 &	0.400 &	1.300 \\
H  & 1969 &	5.086 &	2.130 &	1.415 &	0.775 &	0.425 &	1.475 \\
I  & 1996 &	5.724 &	2.420 &	1.610 &	0.850 &	0.450 &	1.650 \\
J  & 2023 &	6.362 &	2.710 &	1.805 &	0.925 &	0.475 &	1.825 \\
K  & 2050 &	7.000 &	3.000 &	2.000 &	1.000 &	0.500 &	2.000 \\
\hline
\end{tabular}
\end{table}
\subsection{Model performance}
Due to imperfection of the chemical mechanism time evolution of species concentrations given by the model does not completely agree with the photochemical theory, but reproduces it to some extent. Figure \ref{fig:model_validation} shows an example of output from the model, namely the changes in mixing ratios of the main species within 24-hours of simulation time under contrasting $NO_x$ and $VOCs$ conditions. The values shown in Figure \ref{fig:model_validation} have been justified from the theoretical point of view and also compared with those derived from the Atmospheric Chemistry Learner's Box Model (ACLBM, http://guille.uea.ac.uk/ACLBM/). ACLBM is an on-line box model for atmospheric chemistry developed by Prof. Roland von Glasow (University of East Anglia) that allows the selection of several chemical reaction mechanisms, choice of initial concentrations and emission fluxes. For comparison with our modelling data the ACLBM was set up and initialized with values from Tables \ref{tab:setupO3CO}, \ref{tab:setupNOx} and \ref{tab:setupVOCs}, and the '$O_3$-$NO_x$-$CO$-$CH_4$' chemical mechanism was chosen as it is the closest version of the chemical mechanism available in the ACLBM to our model.

The concentration of $NO_x$ is the main factor which determines whether $O_3$ is produced or removed in the atmosphere, and theoretically there are three regimes of $O_3$ production that depend on $NO_x$ levels: at low $NO_x$ levels the regime is characterized by net $O_3$ removal, at intermediate $NO_x$ (also called $NO_x$-sensitive) - by net $O_3$ formation, and at high $NO_x$ - by net $O_3$ formation or removal depending more on $VOCs$ concentration (therefore called $VOC$-sensitive) (e.g., \citep{Fowler2008}). When the prevailing $NO_x$ levels are low, a net $O_3$ removal takes place because there is not enough $NO$ around to react with the peroxy radicals ($HO_2$ and $RO_2$) to produce $NO_2$, which photolysis in conjunction with the reaction (\ref{reac:O3P+O2}) is the main source of $O_3$ in the troposphere. Some additional $O_3$ removal in this regime occurs due to reaction of $HO_2$ radicals with $O_3$ leading to the regeneration of $OH$ radicals as part of an $O_3$ depleting $OH$-$HO_2$ ($HO_x$) cycle. In other words, the following reaction sequence predominates:
\begin{equation}\label{reac:O3+hv}
O_3 + hv \rightarrow O(^1D) + O_2
\end{equation}
\begin{equation}\label{reac:O1D+H2O}
O(^1D) + H_2O \rightarrow OH + OH
\end{equation}
\begin{equation}\label{reac:O1D=O3P}
O(^1D) \xrightarrow{M} O(^3P
\end{equation}
\begin{equation}\label{reac:O3P+O2}
O(^3P) + O_2 \xrightarrow{M} O_3
\end{equation}
\begin{equation}\label{reac:OH+CO}
OH + CO \rightarrow HO_2 + CO_2
\end{equation}
\begin{equation}\label{reac:OH+RH}
OH + RH \rightarrow RO_2 + H_2O
\end{equation}
\begin{equation}\label{reac:HO2+HO2}
HO_2 + HO_2 \rightarrow H_2O_2 + O_2
\end{equation}
\begin{equation}\label{reac:RO2+HO2}
RO_2 + HO_2 \rightarrow ROOH + O_2
\end{equation}
\begin{equation}\label{reac:O3+HO2}
O_3 + HO_2 \rightarrow OH + 2O_2
\end{equation}
At low $NO_x$ regime (for example, 50 ppt) our modelling data agrees well with the theory (Figure \ref{fig:O3valid}). $O_3$ mixing ratio decreases with time, and this decrease is comparable to one derived from ACLBM (by 6 ppb and 4.5 ppb at low and high $VOCs$ concentrations respectively). $NO$ is rapidly (almost instantly) converted to $NO_2$ representing a state equilibrium between the two molecules in the troposphere (Figures \ref{fig:NOvalid}, \ref{fig:NO2valid}). $CO$ is lost due to its reaction with $OH$ (Figure \ref{fig:COvalid}). Mixing ratios of other species in our model are of the order of magnitude as in ACLBM.

\input{diss_valid_vs_time}

When the prevailing $NO_x$ levels are more than 0.5 ppb \citep{Sillman1999}, but less than a certain critical value, reactions (\ref{reac:HO2+NO}) and (\ref{reac:RO2+NO}) represent the dominant reaction pathways for $HO_2$ and $RO_2$ providing a link for catalytic generation of $O_3$:
\begin{equation}\label{reac:HO2+NO}
HO_2 + NO \rightarrow OH + NO_2
\end{equation}
\begin{equation}\label{reac:RO2+NO}
RO_2 + NO \rightarrow RO + NO_2
\end{equation}
\begin{equation}\label{eq:NO2+hv}
NO_2 + hv \rightarrow NO + O(^3P)
\end{equation}
\begin{equation} %same
!!!!! same O(^3P) + O_2 \xrightarrow{M} O_3
\end{equation}
The competition between reactions (\ref{reac:HO2+NO}) and (\ref{reac:HO2+HO2}) for $HO_2$, and reactions (\ref{reac:RO2+NO}) and (\ref{reac:RO2+HO2}) for $RO_2$, allows a greater number of free-radical propagated $O_3$ forming cycles to happen prior to termination (reactions (\ref{reac:HO2+HO2}) and (\ref{reac:RO2+HO2})), therefore resulting in a net $O_3$ formation.

When the prevailing $NO_x$ levels are high, reactions (\ref{reac:HO2+NO}) and (\ref{reac:RO2+NO}) dominate for $HO_2$ and $RO_2$, making the rate of $O_3$ production being controlled by the availability of these radicals. The availability of $OH$ radical is also important at high $NO_x$ levels because $OH$ connects the $O_3$ production rate with rates of its reactions with $CO$ and $RH$ (reactions (\ref{reac:OH+CO}) and (\ref{reac:OH+RH})), therefore making this regime $VOC$-sensitive. Plus at this point the reaction of $OH$ with $NO_2$ to form nitric acid ($HNO_3$) becomes the major termination reaction for free radicals:
\begin{equation}\label{reac:OH+NO2}
OH + NO_2 \xrightarrow{M} HNO_3
\end{equation}
Under these circumstances, the number of free-radical propagated $O_3$ forming cycles decreases with increasing $NO_x$, and the production rate of $O_3$ decreases. However, if the concentrations of $VOCs$ are high, reactions of $OH$ with $CO$ and $RH$ ((\ref{reac:OH+CO}), (\ref{reac:OH+RH})) can compete with reaction (\ref{reac:OH+NO2}), leading to an increase in the $O_3$ production rate.

Since the chemical mechanism of our model does not include the reaction of $HNO_3$ formation, a good representation of $VOC$-sensitive regime by our model is not expected. It also makes it difficult to split $NO_x$-sensitive and $VOC$-sensitive regimes in our model, because this split is determined by the recurrence (size) of the peroxide- and nitric-acid forming reactions \citep{Sillman1999}, while our mechanism lacks one of these reactions. However, needing to evaluate performance of our model under higher $NO_x$ conditions anyway below we consider just a contrasting case of $NO_x$ levels, for example, 5 ppb (Figure \ref{fig:model_validation}).

At higher $NO_x$ regime (5 ppb) $O_3$ mixing ratio in the model increases with time (Figure \ref{fig:O3valid}) which agrees with the photochemical theory. The net increase is by 186 ppb and 171 ppb at low and high $VOCs$ concentrations respectively. $NO$ is rapidly converted to $NO_2$, but slower than at low $NO_x$ levels because of their higher initial concentrations (Figure \ref{fig:NOvalid}, \ref{fig:NO2valid}). Due to a higher efficiency of reaction (\ref{reac:HO2+NO}) $OH$ mixing ratio is higher at higher $NO_x$ levels, resulting in a more pronounced loss of $CO$ (Figure \ref{fig:COvalid}).

The best way to illustrate the central features of the relationship between $O_3$, $NO_x$ and $VOC$ mentioned above is to construct a ozone isopleth plot in a form shown in Figure \ref{fig:Sillman1999}. The isopleth plot shows that the rate of $O_3$ production is a nonlinear function of the $NO_x$ and $VOC$ concentrations. When $VOC$/$NO_x$ ratios are high, the $O_3$ production rate increases with increasing $NO_x$ eventually reaching a local maximum, and decreasing afterwards as $NO_x$ concentrations continue to increase. The local maximum (the 'ridge line') defines the split into $NO_x$-sensitive and $VOC$-sensitive regimes. Under the 'ridge line' the $VOC$/$NO_x$ ratios are high, and the $O_3$ production rate increases with increasing $NO_x$ and almost insensitive to $VOC$. Above the 'ridge line' the $VOC$/$NO_x$ ratios are low, and the $O_3$ production rate increases with increasing $VOC$ and decreases with increasing $NO_x$ \citep{Sillman2013}.

\begin{figure} % Sillman1999withdotsline and netO3rate_withINORG_NOx_1ppb_100ppb
\centering
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=.99\linewidth]{{./pics/Sillman1999_Fig1_adapted_withdotsrectangle}.png}
  \captionof{figure}{Isopleths giving net rate of $O_3$ production (ppb/h, solid lines) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) for mean summer daytime meteorology and clear skies. The solid lines represent production rates of 1, 2.5, 5, 10, 15, 20 and 30 ppb/h. Adapted from \citep{Sillman1999}.}\label{fig:Sillman1999}
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_withINORG_NOx_1ppb_100ppb.png}
  \captionof{figure}{Isopleths giving net rate of $O_3$ production (ppb/h, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) in the first modelling experiment with alkyl nitrate chemistry switched off combined with an additional run with inorganic chemistry only (i.e. the only $VOC$ is $CO$).}\label{fig:noAN_netO3rate}
\end{minipage}
\end{figure}

Rewrite!

The same kind of plot can be found in Figure \ref{fig:noAN_netO3rate}, which is produced using data from our model from the first experiment with alkyl nitrate chemistry switched off. Taking into account the fact that the maximum $NO_x$ mixing ratio considered in the experiment is 10 ppb, our results are directly comparable only with a part of the data from Figure \ref{fig:Sillman1999}, namely with those inside the green dashed square. Moreover, since the minimum $NO_x$ mixing ratio considered in the experiment is 5 ppt, while the corresponding value in Figure \ref{fig:Sillman1999} is 1 ppb, a negative $O_3$ production rate is seen in our results, whereas it stays out of range of Figure \ref{fig:Sillman1999}.

One can see the same pattern in these two figures. At high $VOC$/$NO_x$ ratios, the rate of $O_3$ production increases with increasing $NO_x$ and reaches its maximum at the highest $NO_x$ and $VOC$ levels (which corresponds to the top right corner of the green dashed square). The change in $O_3$ production rate at high $VOC$/$NO_x$ ratios in our model is largely insensitive to $VOC$, which fits into the theory. At low $VOC$/$NO_x$ ratios the $O_3$ production rate starts to depend on $VOC$ concentration also resembling the behaviour of ozone isopleths at the top left corner of the green dashed square. The overall tendency of $O_3$ production rate to increase with increasing $VOC$ and $NO_x$ is present in both figures showing that the model reproduces this feature according to the theory. However, the modelled values of $O_3$ production rates are much lower than those in Figure \ref{fig:Sillman1999}: in the model the maximum production rate is 9 ppb/h, while in theory it is 30 ppb/h, meaning that our model underestimates the $O_3$ production rate, but reproduces some features of the relationship between $O_3$, $NO_x$ and $VOC$.

\section{Results} \label{sec:res}
\subsection{Sensitivity of $O_3$ production to alkyl nitrate chemistry}

To evaluate the sensitivity of $O_3$ production (efficiency) to alkyl nitrate ($RONO_2$) chemistry under different $NO_x$ and $VOC$ conditions a number of simulations were carried out. In these simulations the chemical mechanism of the model was adjusted to either include or exclude the formation and loss of alkyl nitrates, which was accomplished by switching on and off all reactions that involve $RONO_2$ while keeping the rest of the mechanism unchanged.

As it has been noted in Section \ref{sec:intro}, the relationship between $O_3$ and $RONO_2$ production rates is controlled by the branching ratio ($\alpha_3$) between reactions:
\begin{equation}\label{reac:RO2+NO=RO+NO2}
RO_2 + NO \xrightarrow{1-\alpha_3} RO + NO_2
\end{equation}
\begin{equation}\label{reac:RO2+NO=RONO2}
RO_2 + NO \xrightarrow{\alpha_3} RONO_2
\end{equation}
Reaction (\ref{reac:RO2+NO=RO+NO2}) leads to $O_3$ production and $NO$ regeneration, resulting in continuing $O_3$ production, while reaction (\ref{reac:RO2+NO=RONO2}) terminates this process. Branching ratios used in the model are taken from the MCM v3.3, and their values (Table \ref{tab:ANbranching}) were kept constant in all model runs.

\begin{table} %  AN branching ratios
\caption{Branching ratios for the formation of alkyl nitrates from their precursor peroxy radicals and $NO$ (taken from the MCM v3.3 (2015)).}\label{tab:ANbranching}
\centering
\begin{tabular}{cc}
\hline
Alkyl nitrate    & Branching ratio \\
\hline
methyl           & 0.001 \\
ethyl            & 0.009 \\
iso-propyl       & 0.042 \\
n-propyl         & 0.020 \\
n-butyl          & 0.033 \\
sec-butyl        & 0.090 \\
iso-butyl        & 0.033 \\ 
tert-butyl       & 0.025 \\
n-pentyl         & 0.052 \\
2-pentyl         & 0.129 \\
3-pentyl         & 0.131 \\
2-methylbutyl    & 0.052 \\
3-methyl-2-butyl & 0.141 \\
2-methyl-2-butyl & 0.047 \\
\hline
\end{tabular}
\end{table}

Figure \ref{fig:netO3rate_noAN_withAN_diff} shows isopleths of net $O_3$ production rate as a function of $NO_x$ and sum of $VOCs$ initial concentrations from two series of model runs (with and without alkyl chemistry) and the difference between them. Since the actual formation of $O_3$ in our model starts at 250 ppt of $NO_x$, Figure \ref{fig:netO3rate_noAN_withAN_diff} summarises modelling data from experiments with $NO_x$ initial mixing ratios from 250 ppt till 100 ppb.

\begin{figure} % netO3rate noAN withAN diff
\centering
\begin{minipage}{.3\textwidth}
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3rate_noINORG_NOx_250ppt_100ppb.png}
\end{minipage}
\begin{minipage}{.3\textwidth}
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3rate_noINORG_NOx_250ppt_100ppb.png}
\end{minipage}
\begin{minipage}{.3\textwidth}
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3ratediff_noINORG_NOx_250ppt_100ppb.png}
\end{minipage}
\caption{Isopleths giving net rate of $O_3$ production (ppb/h, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle). Difference between them is shown on the right.}\label{fig:netO3rate_noAN_withAN_diff}
\end{figure}

As expected, the inclusion of alkyl nitrate chemistry slows down the net $O_3$ production rate. At low $NO_x$ and $VOCs$ initial levels, this slowdown is -0.2...-0.6 ppb of $O_3$ per hour. As $NO_x$ and $VOCs$ levels increase the $O_3$ production rate decreases, reaching the maximum decrease of -1.94 ppb of $O_3$ per hour at 100 ppb of $NO_x$ and $VOCs$ level from case I (not the highest $VOCs$ level). At the range of $NO_x$ levels from 1 ppb to 10 ppb the $O_3$ production rate is slightly more sensitive to $VOCs$ levels, while it is almost insensitive to $VOCs$ at $NO_x$ levels higher than 10 ppb and 2050 ppbC of $VOCs$.

\begin{figure} % netO3mixrat noAN withAN diff
\centering
\begin{minipage}{.3\textwidth}
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/noAN/chem_noAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.png}
\end{minipage}
\begin{minipage}{.3\textwidth}
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixrat_noINORG_NOx_250ppt_100ppb.png}
\end{minipage}
\begin{minipage}{.3\textwidth}
  \centering
  \includegraphics[width=.99\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_netO3mixratdiff_noINORG_NOx_250ppt_100ppb.png}
\end{minipage}
\caption{Isopleths giving net $O_3$ production (ppb, color) as a function of $VOC$ (ppbC) and $NO_x$ (ppb) with alkyl nitrate chemistry switched off (left) and switched on (middle). Difference between them is shown on the right. Net production = final - initial value of $O_3$ mixing ratio.}\label{fig:netO3mixrat_noAN_withAN_diff}
\end{figure}

The change in net rate of $O_3$ production due to inclusion of alkyl nitrate chemistry into the chemical mechanism has a direct impact on absolute $O_3$ levels calculated by the model. Figure \ref{fig:netO3mixrat_noAN_withAN_diff} shows that the net $O_3$ production (difference between final and initial $O_3$ mixing ratio) is smaller when alkyl nitrate chemistry is switched on. At low $NO_x$ and $VOCs$ levels the (negative) change in the net $O_3$ production is 5-10 ppb, at high $NO_x$ and $VOCs$ levels it is 35-45 ppb (per 24 hours) with maximum value of 46.5 ppb. However, these changes in the net $O_3$ production should be treated with care, because the model due to reasons described in Section \ref{sec:method} gives unrealistically high $O_3$ mixing ratios (maximum 650 ppb while typical maximum urban $O_3$ mixing ratio (Los Angeles, Mexico City) is 490 ppb (http://www-personal.umich.edu/~sillman/ozone.htm). Therefore, in this case a more useful estimate of effect of these changes is to express them in percentage form (i.e., compare like with like), and it appears that these changes are very small: ??? respectively.

To see what caused these changes it is useful to look at the net production of the sum of the alkyl nitrates themselves as a function of $NO_x$ and $VOCs$ levels (Figure \ref{fig:endtotalANs}). Since alkyl nitrates are the only sink for $NO_x$ in our model the high production of these species in total was expected. The range is ppb.
However, real concentrations reported from observational studies are of different order, tens of ppt.
\citep{Reeves2007} 10 ppt (ITOP) \citep{Roberts1998} 14 ppt

\citep{Reeves2007}
[In the 'marine' air masses methyl nitrate shows a slightly higher mean concentration to 'other' air masses (2.40 compared to 2.29 ppt), but the concentrations of the other alkyl
nitrates are lower by around 30–50% with the exception of
ethyl nitrate which is lower by about 15% (1.33 compared to
1.57 ppt), such that it becomes the second most abundant alkyl
nitrate.
The highest alkyl
nitrate concentration was observed during flight B031 (air
from New York) for 2-methyl-3-butyl nitrate (13.7 ppt),
when the maximum concentration of 2 + 3-pentyl nitrate
(9.0 ppt) was also measured. It is quite possible that some of
the ‘‘marine’’ maximum concentrations of the alkyl nitrates
were impacted by terrestrial emissions (e.g., 5.4 pptv of
2-methyl-3-butyl nitrate and 1.7 pptv for 2 + 3-pentyl nitrate
observed during flight B030).

The mean NO concentration observed during ITOP (20 plus minus 31 pptv (Lewis et al., 2007)) is below the limit of 100–200 pptv at which the efficiency of the alkyl nitrate formation is reduced because of competition between the reaction of ROO with NO and ROO self/cross reactions (Roberts et al., 1998)]

\begin{figure} % endtotalANs
\includegraphics[width=.5\linewidth]{D:/FACSIMILE/ANsCBmodel/matlab/ANsCB_pics/ANsNOxVOC/allAN/chem_allAN_endtotalANs_noINORG_NOx_250ppt_100ppb.png}
\caption{Isopleths of net production of alkyl nitrates ($\Sigma RONO_2$) as a function of $NO_x$ and $VOCs$ levels.}\label{fig:endtotalANs}
\end{figure}

! Dependence of alkyl nitrate yield on carbon number (paper about n-alkanes): bigger yield of ANs from larger carbon number means that the potential for contributing to photochemical air pollution may be less for the larger (C>6) n-alkanes that for the smaller ones.

\subsection{$O_3$ production and carbon bonds}

\section{Discussion} \label{sec:discuss}
Future plans and possible improvements in the model.

The experimental results presented in the previous section provide an opportunity to speculate about the behaviour of our simple model of alkyl nitrate production and loss with time in comparison with the observational data.

\section{Conclusion} \label{sec:conclusion}
\bibliography{diss_refs}

\section{Appendix I} \label{sec:appendix1}

\end{document}