### 2020-04-06 code.tex
re-structure msc thesis code

\begin{verbatim}
*-------------------------------REACTIONS------------------------------;
*-------------------------------INORGANIC------------------------------;
*---Inorganic (0_00);
RI1A  % 6.0D-34*O2*(T/300)@-2.6*O2   : O = O3;
RI1B  % 5.6D-34*N2*(T/300)@-2.6*O2   : O = O3;
RI2   % 8.0D-12*EXP(-2060/T)         : O+O3 = ;
RI3   % 3.2D-11*EXP(67/T)*O2         : O1D = O;
RI12A % J<1>                         : O3 = O1D; 
RI12B % J<2>                         : O3 = O;
RI19  % 2.2D-10*H2O                  : O1D = OH+OH;
RI20  % 7.7D-12*EXP(-2100/T)         : OH+H2 = HO2;
RI23  % 4.8D-11*EXP(250/T)           : OH+HO2 = ;
RI24  % 2.9D-12*EXP(-160/T)          : OH+H2O2 = HO2;
RI25  % 1.7D-12*EXP(-940/T)          : OH+O3 = HO2;
RI26  % 2.2D-13*EXP(600/T)*KI2627    : HO2+HO2 = H2O2;
RI27  % 1.9D-33*N2*EXP(980/T)*KI2627 : HO2+HO2 = H2O2;
RI28  % 2.03D-16*(T/300)@4.57*EXP(693/T): HO2+O3 = OH;
RI30  % J<3>                         : H2O2 = OH+OH;
RI36  % 1.8D-11*EXP(107/T)*N2        : O1D = O;
RI45  % 3.6D-12*EXP(270/T)           : HO2+NO = OH+NO2;
RI54  % 1.4D-12*EXP(-1310/T)         : NO+O3 = NO2;
RI68  % J<4>                         : NO2 = NO+O;
RCH42 % 1.44D-13*(1+(M/4.2D+19))     : OH+CO = HO2; 
* RI69  % KMT08                        : OH+NO2 = HNO3;
* RI70  % J<8>                         : HNO3 = OH+NO2;
*--------------------------------ORGANIC-------------------------------;
*---Methane (0_01);
RCH41 % 1.85D-12*EXP(-1690/T)        : OH+CH4 = CH3O2; 
RCH43 % 5.4D-12*EXP(135/T)           : OH+HCHO = HO2+CO; 
RCH44 % 5.3D-12*EXP(190/T)*0.6       : OH+CH3OOH = CH3O2; 
RCH45 % 5.3D-12*EXP(190/T)*0.4       : OH+CH3OOH = HCHO+OH;    
RCH46 % 3.8D-13*EXP(780/T)*(1-1/(1+498*EXP(-1160/T))):HO2+CH3O2=CH3OOH; 
RCH47 % 3.8D-13*EXP(780/T)*(1/(1+498*EXP(-1160/T)))  :HO2+CH3O2=HCHO; 
RCH48 % 7.2D-14*EXP(-1080/T)*O2      : CH3O = HCHO+HO2;
RCH49 % 2.3D-12*EXP(360/T)*0.999     : CH3O2+NO = CH3O+NO2; 
RCH410% J<11>                        : HCHO = CO+HO2+HO2; 
RCH411% J<12>                        : HCHO = CO+H2; 
RCH412% J<41>                        : CH3OOH = CH3O+OH; 
*---Ethane (0_02);
RC2H61 % 6.9D-12*EXP(-1000/T)        : OH+C2H6 = C2H5O2;
RC2H62 % 3.8D-13*EXP(900/T)          : HO2+C2H5O2 = C2H5OOH;
RC2H63 % 2.55D-12*EXP(380/T)*0.991   : C2H5O2+NO = C2H5O+NO2;
RC2H64 % J<41>                       : C2H5OOH = C2H5O+OH;
RC2H65 % 1.90D-12*EXP(190/T)         : OH+C2H5OOH = C2H5O2;
RC2H66 % 8.01D-12                    : OH+C2H5OOH = CH3CHO+OH;
RC2H67 % 2.4D-14*EXP(-325/T)*O2      : C2H5O = CH3CHO+HO2;
RC2H68 % 3.0D-12*EXP(20/T)*0.05      : OH+C2H5OH = C2H5O;
RC2H69 % 3.0D-12*EXP(20/T)*0.90      : OH+C2H5OH = CH3CHO+HO2;
RC2H610% J<13>                       : CH3CHO = CH3O2+HO2+CO;
RC2H611% 4.7D-12*EXP(345/T)*0.95     : OH+CH3CHO = CH3CO3;
RC2H612% 4.7D-12*EXP(345/T)*0.05     : OH+CH3CHO = HCOCH2O2;
*---Propane (0_03);
RC3H81 % 7.6D-12*EXP(-585/T)*0.736   : OH+C3H8 = IC3H7O2;
RC3H82 % 7.6D-12*EXP(-585/T)*0.264   : OH+C3H8 = NC3H7O2;
RC3H83 % KRO2HO2*0.520               : IC3H7O2+HO2 = IC3H7OOH;
RC3H84 % 2.7D-12*EXP(360/T)*0.958    : IC3H7O2+NO = IC3H7O+NO2;
RC3H85 % KRO2HO2*0.520               : NC3H7O2+HO2 = NC3H7OOH;
RC3H86 % 2.9D-12*EXP(350/T)*0.980    : NC3H7O2+NO = NC3H7O+NO2;
RC3H87 % J<41>                       : IC3H7OOH = IC3H7O+OH;
RC3H88 % 1.66D-11                    : OH+IC3H7OOH = CH3COCH3+OH;
RC3H89 % 1.90D-12*EXP(190/T)         : OH+IC3H7OOH = IC3H7O2;
RC3H810% 1.5D-14*EXP(-230/T)*O2      : IC3H7O = CH3COCH3+HO2;
RC3H811% 8.8D-12*EXP(-1320/T)+1.7D-14*EXP(423/T):CH3COCH3+OH=CH3COCH2O2; 
RC3H812% J<21>                       : CH3COCH3 = CH3CO3+CH3O2;
RC3H813% J<41>                       : NC3H7OOH = NC3H7O+OH;
RC3H814% 1.10D-11                    : OH+NC3H7OOH = C2H5CHO+OH;
RC3H815% 1.90D-12*EXP(190/T)         : OH+NC3H7OOH = NC3H7O2;
RC3H816% 2.6D-14*EXP(-255/T)*O2      : NC3H7O = C2H5CHO+HO2;
RC3H817% 4.9D-12*EXP(405/T)          : C2H5CHO+OH = C2H5CO3;
RC3H818% J<14>                       : C2H5CHO = C2H5O2+HO2+CO;
*---n-butane (0_04);
RNC4H101 % 9.8D-12*EXP(-425/T)*0.127 : OH+NC4H10 = NC4H9O2;
RNC4H102 % 9.8D-12*EXP(-425/T)*0.873 : OH+NC4H10 = SC4H9O2;
RNC4H103 % KRO2HO2*0.625             : NC4H9O2+HO2 = NC4H9OOH;
RNC4H104 % KRO2NO*0.967              : NC4H9O2+NO = NC4H9O+NO2;
RNC4H105 % KRO2HO2*0.625             : SC4H9O2+HO2 = SC4H9OOH;
RNC4H106 % KRO2NO*0.910              : SC4H9O2+NO = SC4H9O+NO2;
RNC4H107 % J<41>                     : NC4H9OOH = NC4H9O+OH;	
RNC4H108 % 1.24D-11                  : OH+NC4H9OOH = C3H7CHO+OH;
RNC4H109 % 1.90D-12*EXP(190/T)       : OH+NC4H9OOH = NC4H9O2;
RNC4H1010% 8.9D-14*EXP(-550/T)*O2    : NC4H9O = C3H7CHO+HO2;
RNC4H1011% 4.6D10*EXP(-3570/T)       : NC4H9O = HO1C4O2;
RNC4H1012% J<41>                     : SC4H9OOH = SC4H9O+OH;
RNC4H1013% 5.6D12*EXP(-5780/T)       : SC4H9O = CH3CHO+C2H5O2;
RNC4H1014% 1.5D-14*EXP(-200/T)*O2    : SC4H9O = MEK+HO2 ;
RNC4H1015% J<22>                     : MEK = CH3CO3+C2H5O2;
*---i-butane;
RIC4H101 % 1.16D-17*T@2*EXP(225/T)*0.206 : OH+IC4H10 = IC4H9O2;
RIC4H102 % 1.16D-17*T@2*EXP(225/T)*0.794 : OH+IC4H10 = TC4H9O2;
RIC4H103 % KRO2HO2*0.625                 : IC4H9O2+HO2 = IC4H9OOH;
RIC4H104 % KRO2NO*0.967                  : IC4H9O2+NO = IC4H9O+NO2;
RIC4H105 % KRO2HO2*0.625                 : TC4H9O2+HO2 = TC4H9OOH;
RIC4H106 % KRO2NO*0.975                  : TC4H9O2+NO = TC4H9O+NO2;
RIC4H107 % J<41>                         : IC4H9OOH = IC4H9O+OH;
RIC4H108 % 1.90D-12*EXP(190/T)           : OH+IC4H9OOH = IC4H9O2;
RIC4H109 % 1.24D-11                      : OH+IC4H9OOH = IPRCHO+OH;
RIC4H1010% KROPRIM*O2                    : IC4H9O = IPRCHO+HO2;
RIC4H1011% 2.17D-12*EXP(190/T)           : OH+TC4H9OOH = TC4H9O2;
RIC4H1012% J<41>                         : TC4H9OOH = TC4H9O+OH;
RIC4H1013% 6.00D+14*EXP(-8153/T)         : TC4H9O = CH3COCH3+CH3O2;
*---n-pentane (0_05);
RNC5H121 % 2.44D-17*T@2*EXP(183/T)*0.083 : OH+NC5H12 = PEAO2;
RNC5H122 % 2.44D-17*T@2*EXP(183/T)*0.568 : OH+NC5H12 = PEBO2;
RNC5H123 % 2.44D-17*T@2*EXP(183/T)*0.349 : OH+NC5H12 = PECO2;
RNC5H124 % KRO2HO2*0.706                 : PEAO2+HO2 = PEAOOH;
RNC5H125 % KRO2NO*0.948                  : PEAO2+NO = PEAO+NO2;
RNC5H126 % KRO2HO2*0.706                 : PEBO2+HO2 = PEBOOH;
RNC5H127 % KRO2NO*0.871                  : PEBO2+NO = PEBO+NO2;
RNC5H128 % KRO2HO2*0.706                 : PECO2+HO2 = PECOOH;
RNC5H129 % KRO2NO*0.869                  : PECO2+NO = PECO+NO2;
RNC5H1210% 1.41D-11                      : OH+PEAOOH = C4H9CHO+OH;
RNC5H1211% 1.90D-12*EXP(190/T)           : OH+PEAOOH = PEAO2;
RNC5H1212% J<41>                         : PEAOOH = PEAO+OH;
RNC5H1213% 2.30D-11                      : OH+PEBOOH = MPRK+OH;
RNC5H1214% 1.90D-12*EXP(190/T)           : OH+PEBOOH = PEBO2;
RNC5H1215% J<41>                         : PEBOOH = PEBO+OH;
RNC5H1216% 2.73D-11                      : OH+PECOOH = DIEK+OH;
RNC5H1217% J<41>                         : PECOOH = PECO+OH;
*---i-pentane;
RIC5H121 % 3.70D-12*0.087                : OH+IC5H12 = IPEAO2;
RIC5H122 % 3.70D-12*0.297                : OH+IC5H12 = IPEBO2;
RIC5H123 % 3.70D-12*0.616                : OH+IC5H12 = IPECO2;
RIC5H124 % KRO2HO2*0.706                 : IPEAO2+HO2 = IPEAOOH;
RIC5H125 % KRO2NO*0.948                  : IPEAO2+NO = IPEAO+NO2;
RIC5H126 % KRO2HO2*0.706                 : IPEBO2+HO2 = IPEBOOH;
RIC5H127 % KRO2NO*0.859                  : IPEBO2+NO = IPEBO+NO2;
RIC5H128 % KRO2HO2*0.706                 : IPECO2+HO2 = IPECOOH;
RIC5H129 % KRO2NO*0.953                  : IPECO2+NO = IPECO+NO2;
RIC5H1210% J<41>                         : IPEAOOH = IPEAO+OH;
RIC5H1211% 1.41D-11                      : OH+IPEAOOH = BUT2CHO+OH;
RIC5H1212% 1.90D-12*EXP(190/T)           : OH+IPEAOOH = IPEAO2;
RIC5H1213% J<41>                         : IPEBOOH = IPEBO+OH;
RIC5H1214% 1.90D-12*EXP(190/T)           : OH+IPEBOOH = IPEBO2;
RIC5H1215% 2.29D-11                      : OH+IPEBOOH = MIPK+OH;
RIC5H1216% J<41>                         : IPECOOH = IPECO+OH;
RIC5H1217% 5.25D-12                      : OH+IPECOOH = IPECO2;
*-------------------------------ALKYL NITRATES-------------------------;
*---Methyl nitrate (1_01);
* RCH4NO31 % 2.3D-12*EXP(360/T)*0.001  : CH3O2+NO = CH3NO3;
* RCH4NO32 % 4.0D-13*EXP(-845/T)       : OH+CH3NO3 = HCHO+NO2;
* RCH4NO33 % J<51>                     : CH3NO3 = CH3O+NO2; 
*---Ethyl nitrate (1_02);
* RC2H6NO31 % 2.55D-12*EXP(380/T)*0.009: C2H5O2+NO = C2H5NO3;
* RC2H6NO32 % 6.7D-13*EXP(-395/T)      : OH+C2H5NO3 = CH3CHO+NO2;
* RC2H6NO33 % J<52>                    : C2H5NO3 = C2H5O+NO2;
*---Propyl nitrates (1_03);
* RC3H7NO31 % 2.7D-12*EXP(360/T)*0.042 : IC3H7O2+NO = IC3H7NO3;
* RC3H7NO32 % 6.2D-13*EXP(-230/T)      : OH+IC3H7NO3 = CH3COCH3+NO2;
* RC3H7NO33 % J<54>                    : IC3H7NO3 = IC3H7O+NO2;
* RC3H7NO34 % 2.9D-12*EXP(350/T)*0.020 : NC3H7O2+NO = NC3H7NO3;
* RC3H7NO35 % 5.8D-13                  : OH+NC3H7NO3 = C2H5CHO+NO2;
* RC3H7NO36 % J<53>                    : NC3H7NO3 = NC3H7O+NO2;
*---n-butyl nitrates (1_04);
* RNC4H10N1 % KRO2NO*0.033             : NC4H9O2+NO = NC4H9NO3;
* RNC4H10N2 % KRO2NO*0.090             : SC4H9O2+NO = SC4H9NO3;
* RNC4H10N3 % 1.6D-12                  : OH+NC4H9NO3 = C3H7CHO+NO2;
* RNC4H10N4 % 8.6D-13                  : OH+SC4H9NO3 = MEK+NO2;
* RNC4H10N5 % J<53>                    : NC4H9NO3 = NC4H9O+NO2;
* RNC4H10N6 % J<54>                    : SC4H9NO3 = SC4H9O+NO2;
*---i-butyl nitrates;
* RIC4H10N1 % KRO2NO*0.033             : IC4H9O2+NO = IC4H9NO3;
* RIC4H10N2 % KRO2NO*0.025             : TC4H9O2+NO = TC4H9NO3;
* RIC4H10N3 % 7.69D-13                 : OH+IC4H9NO3 = IPRCHO+NO2;
* RIC4H10N4 % 8.16D-14                 : OH+TC4H9NO3 = CH3COCH3+HCHO+NO2;
* RIC4H10N5 % J<53>                    : IC4H9NO3 = IC4H9O+NO2;
* RIC4H10N6 % J<55>                    : TC4H9NO3 = TC4H9O+NO2;
*---n-pentyl nitrates (1_05);
* RNC5H12N1 % KRO2NO*0.052             : PEAO2+NO = PEANO3;
* RNC5H12N2 % KRO2NO*0.129             : PEBO2+NO = PEBNO3;
* RNC5H12N3 % KRO2NO*0.131             : PECO2+NO = PECNO3;
* RNC5H12N4 % 3.00D-12                 : OH+PEANO3 = C4H9CHO+NO2;
* RNC5H12N5 % 1.85D-12                 : OH+PEBNO3 = MPRK+NO2;
* RNC5H12N6 % 1.12D-12                 : OH+PECNO3 = DIEK+NO2;
* RNC5H12N7 % J<53>                    : PEANO3 = PEAO+NO2;
* RNC5H12N8 % J<54>                    : PEBNO3 = PEBO+NO2;
* RNC5H12N9 % J<54>                    : PECNO3 = PECO+NO2;
*---i-pentyl nitrates;
RIC5H12N1 % KRO2NO*0.052             : IPEAO2+NO = IPEANO3;
RIC5H12N2 % KRO2NO*0.141             : IPEBO2+NO = IPEBNO3;
RIC5H12N3 % KRO2NO*0.047             : IPECO2+NO = IPECNO3;
RIC5H12N4 % 2.01D-12                 : OH+IPEANO3 = BUT2CHO+NO2;
RIC5H12N5 % 1.82D-12                 : OH+IPEBNO3 = MIPK+NO2;
RIC5H12N6 % 4.08D-13                 : OH+IPECNO3 = CH3COCH3+CH3CHO+NO2;
RIC5H12N7 % J<53>                    : IPEANO3 = IPEAO+NO2;
RIC5H12N8 % J<54>                    : IPEBNO3 = IPEBO+NO2;
RIC5H12N9 % J<55>                    : IPECNO3 = IPECO+NO2;
\end{verbatim}